---
title: "CVD Incidence - CASE IDENTIFICATION"
author: "Nathalie Conrad"
date: "October 2023"
output: html_document
---

# LIBRARIES AND SETTINGS
```{r, warning=FALSE, message=FALSE}
#install.packages("gdata")
library(data.table)
library(knitr)
#library(gdata)
#library(ggplot2)
library(rapportools)
#library(rapport)
library(bit64)
library(epiR)
library(gdata)
```

# STUDY CRITERIA 

## Sensitivity analyses
```{r, eval=TRUE, warning=FALSE, message=FALSE}
include_cvd_death = "yes" # select yes if fatal CVD events are to be included, select no if only cvd diagnoses (in primary or secondary care are to be included)
# includes cvd deaths recorded in any of the 15 death causes
```

## Study Dates
```{r, eval=TRUE, warning=FALSE, message=FALSE}
#Data: CPRD Jan 2021 Release DR-8944
#study_start_date = as.Date("01/01/1995", format="%d/%m/%Y")
study_start_date = as.Date("01/01/1998", format="%d/%m/%Y")
study_end_date = as.Date("30/06/2019", format="%d/%m/%Y")
```

## Work Environment
```{r, warning=FALSE, message=FALSE}
# Secure data directories
PATH_D = "/Users/u0133446/CPRD-2020/CPRD-DATA-1stRelease" # path to data directory
PATH_D2 = "/Users/u0133446/CPRD-2020/Final_Delivery1_19_156" # path to data directory
PATH_W = "/Users/u0133446/CPRD-2020/Workspace/CVD_incidence" # path to workspace directory
PATH_W_orig = "/Users/u0133446/CPRD-2020/Workspace" # path to workspace directory
#PATH_L = "/Users/u0133446/CPRD-2020/CPRD-DATA-1stRelease/Lookups" # path to lookups directory
PATH_L = "/Users/u0133446/CPRD-2020/Final_Delivery1_19_156/Lookups" # path to lookups directory
#load(file.path(PATH_W,"2022-02-18SelectedWorkspaceIdentification.RData"))

# Dropbox directories for diagnostic codes, aggregated tables and plots
PATH = "/Users/u0133446/Dropbox/Leuven/INF-CVD" # path to dropbox directory
PATH_P = "/Users/u0133446/Dropbox/Leuven/INF-CVD/CPRD-2020_non_data/Plots/CVD_incidence" # path to plots directory
```


## My functions
```{r, eval=TRUE, warning=FALSE, message=FALSE}
compute_days = function (year, data, diag_date){
  year_start = as.Date(paste("01/01/",year,sep=""), format="%d/%m/%Y")
  year_end = as.Date(paste("31/12/",year,sep=""), format="%d/%m/%Y")
  days=as.numeric(pmax(pmin(diag_date, data$end_date,year_end,na.rm=T)+1 - pmax(data$start_date, year_start),0))
return(days)}
#as.numeric(pmax(pmin(diag_date, data$end_date,na.rm=T)+1 - data$start_date,0))
#as.numeric(pmax(pmin(2000-04-14, 2000-04-14,na.rm=T)+1 - 2000-01-01,0))
#as.numeric(pmax(as.Date(14/04/2000, format="%d/%m/%Y")+1 - as.Date(01/01/2000, format="%d/%m/%Y"),0))
#test=as.Date(14-04-2000, format="%d/%m/%Y")

age_group_fun = function (yob, ref_year) { 
  delta = as.numeric(as.character(ref_year)) - yob
  age_group_id = pmin((delta%/%5)+pmin(delta,1),20)
  return (age_group_id)}
```

## Disease codes - (code review 03/03/2023)
```{r disease_codes, eval=TRUE, warning=FALSE, message=FALSE}
# all codes
all_codes_table = data.table(read.table(file.path(PATH,"/Phenotyping/Final_AD_INF_CVD/CVD_incidence/2023-03-03_FINAL_CVD_diagnostic_codes_table_JMM.txt"), header = TRUE, sep="\t", colClasses=c("medcode_aurum"="integer64"))) # Final set of codes for analysis as per JMM review 03/03/2023
nrow(all_codes_table)
head(all_codes_table)
levels(as.factor(all_codes_table$include))

all_codes_table = all_codes_table[include == 1]

head(all_codes_table)
colnames(all_codes_table)
nrow(all_codes_table)
levels(as.factor(all_codes_table$include))
levels(as.factor(all_codes_table$codetype))
levels(as.factor(all_codes_table$disease))

# gold 
gold_codes_table = all_codes_table[codetype == "gold"]
nrow(gold_codes_table)

# aurum
aurum_codes_table = all_codes_table[codetype == "aurum"]
nrow(aurum_codes_table)
# Check that aurum codes have not been shortened
class(aurum_codes_table$medcode_aurum)
summary(aurum_codes_table$medcode_aurum)

#max(as.integer64(diag_read_codes_table1$medcode_aurum))
# if max = 1..e+16 then codes have been truncated
aurum_codes_table[medcode_aurum == "11920121000006117"]
#Transient cerebral ischemia

# icd
icd_codes_table = all_codes_table[codetype == "icd"]
icd_codes_table[,alt_icd_code:= gsub("\\.", "", icd_code)] # remove "." in ICD-10 codes
nrow(icd_codes_table)
icd_codes_table

# icd version 9 and 10 (only used in death records up to the year 2000)
icd_9_10_codes_table = all_codes_table[codetype %in% c("icd9", "icd")]
icd_9_10_codes_table[,alt_icd_code:= gsub("\\.", "", icd_code)] # remove "." in ICD-10 codes
nrow(icd_9_10_codes_table)
icd_9_10_codes_table[codetype == "icd9"]

# opcs
opcs_codes_table= all_codes_table[codetype == "opcs"]
opcs_codes_table[,alt_opcs_code:= gsub("\\.", "", opcs_code)] # remove "." in ICD-10 codes
nrow(opcs_codes_table)

nrow(gold_codes_table)+nrow(aurum_codes_table)+nrow(icd_codes_table)+nrow(opcs_codes_table)

# Full original dictionary including codes labeled as include ==  0-possible (ie to be excluded) - for checks only
diag_codes_table_inc_0 = data.table(read.table(file.path(PATH,"/Phenotyping/Final_AD_INF_CVD/CVD_incidence/2023-03-03_FULL_CVD_diagnostic_codes_table_JMM.txt"), header = TRUE, sep="\t", colClasses=c("medcode_aurum"="integer64"))) # This table includes all codes considered for this analysis including those labeled as include %in% c("0", "0-possible") ONLY for CHECKs
nrow(diag_codes_table_inc_0)
```

# READ MAIN TABLES

## 1. DENOMINATOR FILES

## 1.1 DENOMINATOR in CPRD GOLD
```{r, eval=TRUE, warning=FALSE, message=FALSE}
gold_source_main= fread(file.path(PATH_D,"/CPRD_GOLD/Source/GOLD_interim.txt.gz"), header = T, sep="\t")
gold_source_imd= fread(file.path(PATH_D,"/CPRD_GOLD/Source/GOLD_patient_imd2015_interim.txt.gz"), header = T, sep="\t")
gold_source_ons= fread(file.path(PATH_D,"/CPRD_GOLD/Source/GOLD_ONS_interim.txt.gz"), header = T, sep="\t")

# Check total number of records
nrow(gold_source_main)
nrow(gold_source_imd)
nrow(gold_source_ons)

#summary(gold_source_imd$imd2015_5)

# Check only patients with HES, ONS and SES linkage and ACCEPTable patients are included
nrow(gold_source_main[hes_e==0|death_e==0|lsoa_e==0|accept==0])

# Merge main table with IMD and DEATH tables
denominator_gold = merge(gold_source_main, gold_source_imd[,.(patid, imd2015_5)], by= "patid", all = TRUE)
denominator_gold = merge(denominator_gold, gold_source_ons[,.(patid, dod)], by= "patid", all = TRUE)
colnames(denominator_gold)
#gold_source = merge(gold_source_0, gold_source_ons, by= "patid", all = TRUE)
#gold_source_imd[patid==110010] # check missing data for imd
#nrow(gold_source [!(dod.x == dod.y)])  # check differences in dod data

# Get sex description
gender_lookup= fread(file.path(PATH_L,"/GOLD/TXTFILES/SEX.txt"), header = T, sep="\t")
setnames(gender_lookup, old=c("Code", "Sex"), new =c("gender", "gender_name"))
denominator_gold = merge(denominator_gold, gender_lookup, by= "gender", all.x = TRUE)

# Get Region description
region_lookup= fread(file.path(PATH_L,"/GOLD/TXTFILES/PRG.txt"), header = T, sep="\t")
setnames(region_lookup, old=c("Code", "Practice Region"), new =c("region", "region_name"))
region_lookup[region_name == "Yorkshire & The Humber"]$region_name = "Yorkshire And The Humber"
region_lookup
denominator_gold = merge(denominator_gold, region_lookup, by= "region", all.x = TRUE)

# Check dates
nrow(denominator_gold[is.na(regdate)])
nrow(denominator_gold[is.na(todate)])
nrow(denominator_gold[is.na(lcdate)])

# Set correct date format
denominator_gold[,regdate:=as.Date(regdate, format="%d/%m/%Y")] #set date format
#denominator_gold[,utsdate:=as.Date(utsdate, format="%d/%m/%Y")] #set date format  --- UTS MISSING in 2nd delivery
denominator_gold[,todate:=as.Date(todate, format="%d/%m/%Y")] #set date format
denominator_gold[,dod:=as.Date(dod, format="%d/%m/%Y")] #set date format
denominator_gold[,deathdate:=as.Date(deathdate, format="%d/%m/%Y")] #set date format
denominator_gold[,lcdate:=as.Date(lcdate, format="%d/%m/%Y")] #set date format

# Registration start and end dates
denominator_gold[,birth_date := as.Date(paste("30/06/",birthyear, sep=""), format="%d/%m/%Y")]
#denominator_gold$start_date = pmax(denominator_gold$regdate, study_start_date,denominator_gold$birth_date, denominator_gold$utsdate, na.rm=T)
denominator_gold[,start_date := pmax(regdate, study_start_date, birth_date, na.rm=T)]
denominator_gold[,end_date := pmin(todate, study_end_date, dod, lcdate, na.rm=T)]
denominator_gold[,reg_time:=end_date-start_date]
nrow(denominator_gold[is.na(start_date)])
nrow(denominator_gold[is.na(end_date)])

# Deduplicate practices from GOLD that have moved to Aurum
Vision_to_Emis_practices= fread(file.path(PATH_D,"/CPRD_GOLD/VisionToEmisMigrators_old_format.txt.gz"), header = T, sep="\t")
nrow(Vision_to_Emis_practices)
Vision_to_Emis_practices = unique(Vision_to_Emis_practices)
nrow(Vision_to_Emis_practices)
length(unique(Vision_to_Emis_practices$gold_pracid))
Vision_to_Emis_practices$exclude=1 ## label patids with exclude == 1 => those patients who are both in CPRD gold and aurum
denominator_gold = merge (denominator_gold, Vision_to_Emis_practices, by.x="pracid", by.y="gold_pracid", all.x =TRUE)
nrow(denominator_gold[is.na(exclude)])
denominator_gold[is.na(exclude), exclude:= 0]
nrow(denominator_gold)
head(denominator_gold)

# !! NEW on 15 Sep 2021 - NEW GOLD patids
denominator_gold[,patidold:= patid]
denominator_gold[,pracidold:= pracid]
# change pracids
denominator_gold[,pracid := as.character(substr(patid, floor(log10(patid))-1, floor(log10(patid))+1 ))] # pracid = last three digits of patidold
denominator_gold[,pracid:= paste("10",pracid,sep="")] # new pracid = "10" + last three digits of patidold 
# change patids
denominator_gold[,patid := as.character(substr(patidold, 1, floor(log10(patidold))-2 ))] # patid all digits up to the last three (which are the pracid)
denominator_gold[,patid:= as.numeric(paste(patid,pracid,sep=""))] # new patid = old patid (up to the last three digits) + new pracid 
#denominator_gold[patid == 1002210231]
#1000010010	10010	10000010	10	
#1002210231	10231	10022231	231

## get table with old/new patids
#old_new_patids= fread("/Users/u0133446/CPRD-2020/Final_Delivery1_19_156/GOLD/Data/old_new_patids_pracids.txt.gz") 
#nrow(old_new_patids) # this tables contains new patids only for patids in the second delivery

```

## 1.2 DENOMINATOR in CPRD AURUM
```{r, eval=TRUE, warning=FALSE, message=FALSE}
aurum_source_main= fread(file.path(PATH_D,"/CPRD_Aurum/Source/Aurum_interim.txt.gz"), header = T, sep="\t")
aurum_source_imd= fread(file.path(PATH_D,"/CPRD_Aurum/Source/Aurum_patient_imd2015_interim.txt.gz"), header = T, sep="\t")
aurum_source_ons= fread(file.path(PATH_D,"/CPRD_Aurum/Source/Aurum_ONS_interim.txt.gz"), header = T, sep="\t")

# Check total number of records
nrow(aurum_source_main)
nrow(aurum_source_imd)
nrow(aurum_source_ons)
summary(aurum_source_imd$imd2015_5)

# Check only patients with HES, ONS and SES linkage and ACCEPTable patients are included
nrow(aurum_source_main[hes_e==0|death_e==0|lsoa_e==0|acceptable==0])

# Merge main table with IMD and DEATH tables
denominator_aurum = merge(aurum_source_main, aurum_source_imd[,.(patid, imd2015_5)], by= "patid", all = TRUE)
denominator_aurum = merge(denominator_aurum, aurum_source_ons[,.(patid, dod)], by= "patid", all = TRUE)
setnames(denominator_aurum, old=c("acceptable", "lcd", "yob"), new =c("accept", "lcdate", "birthyear"))

# Get Gender description
denominator_aurum$gender_name = ""
denominator_aurum[gender == "F"]$gender_name = "Female"
denominator_aurum[gender == "M"]$gender_name = "Male"

# Get Region description
denominator_aurum[pracid == 20957]$region = 5
region_lookup_aurum= fread(file.path(PATH_L,"/Aurum/Region.txt"), header = T, sep="\t")
setnames(region_lookup_aurum, old=c("regionid", "Description"), new =c("region", "region_name"))
nrow(denominator_aurum[is.na(region)]) 
denominator_aurum = merge(denominator_aurum, region_lookup_aurum, by= "region", all.x = TRUE)

# Set correct date format
denominator_aurum[,regdate:=as.Date(regstartdate, format="%d/%m/%Y")] #set date format
denominator_aurum[,todate:=as.Date(regenddate, format="%d/%m/%Y")] #set date format
denominator_aurum[,dod:=as.Date(dod, format="%d/%m/%Y")] #set date format
denominator_aurum[,lcdate:=as.Date(lcdate, format="%d/%m/%Y")] #set date format

# Check dates
nrow(denominator_aurum[is.na(regdate)])
nrow(denominator_aurum[is.na(todate)])
nrow(denominator_aurum[is.na(lcdate)])

# Registration start and end dates
denominator_aurum[,birth_date := as.Date(paste("30/06/",denominator_aurum$birthyear, sep=""), format="%d/%m/%Y")]
denominator_aurum[,start_date := pmax(regdate, study_start_date,birth_date, na.rm=T)]
denominator_aurum[,end_date := pmin(todate, study_end_date, dod, lcdate, na.rm=T)]
denominator_aurum[,reg_time:=end_date-start_date]
nrow(denominator_aurum[is.na(start_date)])
nrow(denominator_aurum[is.na(end_date)])
```

## 1.3 Rbind GOLD and AURUM denominator files
```{r, eval=TRUE, warning=FALSE, message=FALSE}
denominator_gold[,gold_db := 1] 
denominator_gold[,db_patid := paste("G", patid, sep="")] 
denominator_aurum[,aurum_db := 1] 
denominator_aurum[,db_patid := paste("A", patid, sep="")] 
denominator = rbind(denominator_gold[exclude==0], denominator_aurum, fill=TRUE) ## exclude patids with exclude == 1 - those patients who are both in CPRD gold and aurum
head(denominator)
nrow(denominator_gold[exclude==0])
nrow(denominator_aurum)
nrow(denominator_gold[exclude==0])+nrow(denominator_aurum)
nrow(denominator)
denominator[is.na(gold_db), gold_db:= 0]
denominator[is.na(aurum_db), aurum_db:= 0]
#rm(denominator_gold, denominator_aurum, aurum_source_main, aurum_source_imd, aurum_source_ons, gold_source_main, gold_source_imd, gold_source_ons)

fwrite(denominator, file.path(PATH_W, "denominator_first_release.txt.gz"))

```

## 1.4 Denominator characteristics
```{r, eval=TRUE, warning=FALSE, message=FALSE}
# Included in the study were men and women with records labelled as ‘acceptable', approved for HES and ONS linkage, and registered with their general practice for at least 12 months during the study period (01/01/ 1998 to 30/06/2019). 
#denominator = fread(file.path(PATH_W, "denominator_first_release.txt.gz"))
format(nrow(denominator), big.mark="'")  # number of records in table
#  22'441'096
colnames(denominator)

# Check only patients with HES, ONS and SES linkage and ACCEPTable patients are included
format(nrow(denominator[hes_e ==1 & death_e==1 & lsoa_e==1 &  accept==1]), big.mark="'")  
denominator
#  22'441'096

# Check patients registered with their general practice for at least 12 months during the study period (01/01/ 1998 to 30/06/2019). 
summary(denominator$reg_time)
format(nrow(denominator[as.numeric(reg_time)>364]), big.mark="'")
denominator[as.numeric(reg_time)<364]
#  21'622'062 = total number of patients included in the study
#22'441'096-21'622'062 = 819'034
```


## 2. EVENTS of INTEREST

## 2.1 Diagnoses in CPRD GOLD

### Diagnoses in CPRD GOLD - Primary Care
```{r, eval=TRUE, warning=FALSE, message=FALSE}
setwd(file.path(PATH_D2, "GOLD/Data/Primary_Care/Clinical"))
#gold_diagnoses_primary_care= fread(file.path(PATH_D2,"/GOLD/Data/Primary_Care/Clinical/19_156_Extract_Clinical_001.txt.gz"), header = T, sep="\t")
#subset(gold_diagnoses_primary_care, medcode %in% gold_codes_table$medcode_gold)
gold_diagnoses_primary_care = rbindlist( lapply(list.files(), function(x) {subset(fread(x), medcode %in% gold_codes_table$medcode_gold)} ))

format(nrow(gold_diagnoses_primary_care), big.mark="'")  # Check total number of events
gold_diagnoses_primary_care[,eventdate:=as.Date(eventdate, format="%d/%m/%Y")] # Set date format
nrow(gold_diagnoses_primary_care[is.na(eventdate),]) # Check number of corrupt eventdates
gold_diagnoses_primary_care=gold_diagnoses_primary_care[!is.na(eventdate),] # Exclude events with corrupt date
nrow(gold_diagnoses_primary_care)
gold_diagnoses_primary_care = gold_diagnoses_primary_care[duplicated(gold_diagnoses_primary_care)==F,] # Exclude duplicates
nrow(gold_diagnoses_primary_care)
gold_diagnoses_primary_care[,database := "CPRD GOLD"] 
gold_diagnoses_primary_care[,codetype := "gold"] 
head(gold_diagnoses_primary_care)

#Merge with code dictionary
gold_diagnoses_primary_care_dict = merge(gold_diagnoses_primary_care[,.(patid, eventdate, medcode, database, codetype)], gold_codes_table[,.(medcode_gold, unique_ID, description, disease, disease_category, disease_subcategory, disease_short, include)], by.x = "medcode", by.y = "medcode_gold", all.x=T, allow.cartesian=TRUE)
setnames(gold_diagnoses_primary_care_dict, old=c("medcode", "unique_ID"), new =c("diag_code", "diag_code_unique_ID")) # rename to diag_code
nrow(gold_diagnoses_primary_care)
nrow(gold_diagnoses_primary_care_dict)
nrow(gold_diagnoses_primary_care_dict[is.na(disease)])
levels(as.factor(gold_diagnoses_primary_care_dict[is.na(disease)]$medcode)) # Codes checked (these are codes that have been excluded from final analyses) 

# check which codes remain invalid
gold_codes_to_be_double_checked_for_inclusion = gold_diagnoses_primary_care_dict[is.na(disease)]
gold_codes_to_be_double_checked_for_inclusion[, count_na:= .N, by = c("diag_code")]
gold_codes_to_be_double_checked_for_inclusion = unique(gold_codes_to_be_double_checked_for_inclusion[,.(diag_code, count_na)])
gold_codes_to_be_double_checked_for_inclusion = merge(gold_codes_to_be_double_checked_for_inclusion, diag_codes_table_inc_0[codetype == "gold",.(description, disease, disease_category, unique_ID, medcode_gold, codetype, include)], by.x = "diag_code", by.y = "medcode_gold", all.x = TRUE) # merge with full original dictionary to retrieve descriptions
gold_codes_to_be_double_checked_for_inclusion = gold_codes_to_be_double_checked_for_inclusion[order(-count_na)]
gold_codes_to_be_double_checked_for_inclusion

# exclude events that refer to codes that are not part of final set of codes
head(gold_diagnoses_primary_care_dict[is.na(disease)])
gold_diagnoses_primary_care_dict = gold_diagnoses_primary_care_dict[!is.na(disease)] # exclude events that refer to codes that are not part of final set of codes
format(nrow(gold_diagnoses_primary_care_dict), big.mark="'")  # number of records in table
head(gold_diagnoses_primary_care_dict)
```

### Diagnoses in CPRD GOLD - HES APC
```{r, eval=TRUE, warning=FALSE, message=FALSE}
# HES APC ICD - hospitalisations
gold_diagnoses_HES_APC_ICD= fread(file.path(PATH_D2,"/GOLD/Data/Linked_Data/HES_Diagnoses_by_hosp/hes_diagnosis_hosp_19_156A_DM.txt.gz"), header = T, sep="\t")
head(gold_diagnoses_HES_APC_ICD)
nrow(gold_diagnoses_HES_APC_ICD)
colnames(gold_diagnoses_HES_APC_ICD)
gold_diagnoses_HES_APC_ICD[as.numeric(icd)>0] # double-check if icd-9 codes were used (icd-9 codes are numeric and should not have been used here)

gold_diagnoses_HES_APC_ICD[,database := "HES APC GOLD"] 
setnames(gold_diagnoses_HES_APC_ICD, old=c("discharged", "icd", "icdx"), new =c("eventdate", "ICD_main", "ICD_modifier")) # date of diagnosis = date of admission 
gold_diagnoses_HES_APC_ICD[,eventdate:=as.Date(eventdate, format="%d/%m/%Y")] #set date format
nrow(gold_diagnoses_HES_APC_ICD[is.na(eventdate),]) # double check number of corrupt eventdates
gold_diagnoses_HES_APC_ICD[is.na(eventdate), eventdate:= as.Date(admidate, format="%d/%m/%Y")]# if discharge date is missing, then use epiend date
#gold_diagnoses_HES_APC_ICD[is.na(eventdate), eventdate:= as.Date(epistart, format="%d/%m/%Y")]# if discharge date is missing, then use epistart date
nrow(gold_diagnoses_HES_APC_ICD[is.na(eventdate),]) # double check number of corrupt eventdates
gold_diagnoses_HES_APC_ICD = gold_diagnoses_HES_APC_ICD[!is.na(eventdate),] # Exclude corrupt eventdates
nrow(gold_diagnoses_HES_APC_ICD)
#nrow(gold_diagnoses_HES_APC_ICD[rapportools::is.empty(eventdate),]) # double check number of empty eventdates
gold_diagnoses_HES_APC_ICD = gold_diagnoses_HES_APC_ICD[duplicated(gold_diagnoses_HES_APC_ICD)==F,] # Exclude duplicates
nrow(gold_diagnoses_HES_APC_ICD)
gold_diagnoses_HES_APC_ICD[,codetype := "icd"] 
#gold_diagnoses_HES_APC[is.empty(ICD_main)&is.empty(ICD_modifier)]$codetype = "opcs"
gold_diagnoses_HES_APC_ICD[ICD_modifier %in% c("-","--","~","-~", "A", "D", "d", "Z", "-A","-D", "X", "E", "+", "*", ",", "!", "G", ".", "H", "F", "J", "K", "R", "N", "DN", "C", "I", "DH", "DR", "H3", "+L", "DG", "a", "M", "#", "~Y", "S", "0D", "2D", "2~", "9~","9D")]$ICD_modifier = ""  # Exclude unhelpful modifiers
gold_diagnoses_HES_APC_ICD[,ICD := paste(ICD_main, ICD_modifier, sep="")] # concatenate ICD codes and modifiers as one single code
gold_diagnoses_HES_APC_ICD = unique(gold_diagnoses_HES_APC_ICD[,.(patid, eventdate, ICD, database, codetype)]) # Exclude duplicates
nrow(gold_diagnoses_HES_APC_ICD)

# HES APC OPCS
gold_diagnoses_HES_APC_OPCS= fread(file.path(PATH_D2,"/GOLD/Data/Linked_Data/HES_Procedures/hes_procedures_epi_19_156A_DM.txt.gz"), header = T, sep="\t")
head(gold_diagnoses_HES_APC_OPCS)
nrow(gold_diagnoses_HES_APC_OPCS)
colnames(gold_diagnoses_HES_APC_OPCS)

gold_diagnoses_HES_APC_OPCS[,database := "HES APC GOLD"] 
setnames(gold_diagnoses_HES_APC_OPCS, old=c("discharged"), new =c("eventdate")) # date of diagnosis = date of admission 
gold_diagnoses_HES_APC_OPCS[,eventdate:=as.Date(eventdate, format="%d/%m/%Y")] #set date format
nrow(gold_diagnoses_HES_APC_OPCS[is.na(eventdate),]) # double check number of corrupt eventdates
gold_diagnoses_HES_APC_OPCS[is.na(eventdate), eventdate:= as.Date(epiend, format="%d/%m/%Y")]# if discharge date is missing, then use epiend date
gold_diagnoses_HES_APC_OPCS[is.na(eventdate), eventdate:= as.Date(epistart, format="%d/%m/%Y")]# if discharge date is missing, then use epistart date
nrow(gold_diagnoses_HES_APC_OPCS[is.na(eventdate),]) # double check number of corrupt eventdates
gold_diagnoses_HES_APC_OPCS = gold_diagnoses_HES_APC_OPCS[!is.na(eventdate),] # Exclude corrupt eventdates
nrow(gold_diagnoses_HES_APC_OPCS)
#nrow(gold_diagnoses_HES_APC_OPCS[rapportools::is.empty(eventdate),]) # double check number of empty eventdates
gold_diagnoses_HES_APC_OPCS = gold_diagnoses_HES_APC_OPCS[duplicated(gold_diagnoses_HES_APC_OPCS)==F,] # Exclude duplicates
nrow(gold_diagnoses_HES_APC_OPCS)
gold_diagnoses_HES_APC_OPCS[,codetype := "opcs"] 
gold_diagnoses_HES_APC_OPCS = unique(gold_diagnoses_HES_APC_OPCS[,.(patid, eventdate, opcs, database, codetype)]) # Exclude duplicates
nrow(gold_diagnoses_HES_APC_OPCS)


#ICD table - Merge with code dictionary
#icd_codes_table
gold_diagnoses_HES_APC_ICD_dict = merge(gold_diagnoses_HES_APC_ICD, icd_codes_table[,.(icd_code, alt_icd_code, unique_ID, description, disease, disease_category,disease_subcategory, disease_short, include)], by.x = "ICD", by.y = "icd_code", all.x=T, allow.cartesian=TRUE)
setnames(gold_diagnoses_HES_APC_ICD_dict, old=c("ICD", "unique_ID"), new =c("diag_code", "diag_code_unique_ID")) # rename columns
nrow(gold_diagnoses_HES_APC_ICD)
nrow(gold_diagnoses_HES_APC_ICD_dict)
nrow(gold_diagnoses_HES_APC_ICD_dict[is.na(disease)])
levels(as.factor(gold_diagnoses_HES_APC_ICD_dict[is.na(disease)]$ICD)) # Codes checked (these are codes that have been excluded from final analyses) 
nrow(gold_diagnoses_HES_APC_ICD_dict)
head(gold_diagnoses_HES_APC_ICD_dict)

# for codes that do not match, try to match with removing last character
gold_diagnoses_HES_APC_ICD_dict_na = gold_diagnoses_HES_APC_ICD_dict[is.na(disease)]
nrow(gold_diagnoses_HES_APC_ICD_dict_na)
#gold_diagnoses_HES_APC_ICD_dict_na[,orig_diag_code := diag_code]
gold_diagnoses_HES_APC_ICD_dict_na[,short_diag_code := substr(diag_code, 1, nchar(diag_code)-1)]
#gold_diagnoses_HES_APC_ICD_dict[is.na(disease), diag_code:= mod_diag_code]
gold_diagnoses_HES_APC_ICD_dict_na = merge(unique(gold_diagnoses_HES_APC_ICD_dict_na[,.(patid, eventdate, diag_code, short_diag_code, database, codetype)]), icd_codes_table[,.(icd_code, unique_ID, description, disease, disease_category,disease_subcategory, disease_short, include)], by.x = "short_diag_code", by.y = "icd_code", all.x=T, allow.cartesian=TRUE)
head(gold_diagnoses_HES_APC_ICD_dict_na[!is.na(disease)])
gold_diagnoses_HES_APC_ICD_dict_na[,diag_code_orig:=diag_code]
gold_diagnoses_HES_APC_ICD_dict_na[,diag_code:=short_diag_code]
gold_diagnoses_HES_APC_ICD_dict_na[,diag_code_unique_ID:=unique_ID]
nrow(gold_diagnoses_HES_APC_ICD_dict_na)
nrow(gold_diagnoses_HES_APC_ICD_dict_na[is.na(disease)])
gold_diagnoses_HES_APC_ICD_dict_na[is.na(disease)]

# check which codes remain invalid
icd_codes_to_be_double_checked_for_inclusion_gold = gold_diagnoses_HES_APC_ICD_dict_na[is.na(disease)]
icd_codes_to_be_double_checked_for_inclusion_gold[, count_na:= .N, by = c("diag_code")]
icd_codes_to_be_double_checked_for_inclusion_gold = unique(icd_codes_to_be_double_checked_for_inclusion_gold[,.(diag_code, short_diag_code, count_na)])
icd_codes_to_be_double_checked_for_inclusion_gold = merge(icd_codes_to_be_double_checked_for_inclusion_gold, diag_codes_table_inc_0[codetype == "icd",.(description, disease, disease_category, unique_ID, icd_code,codetype, include)], by.x = "diag_code", by.y = "icd_code", all.x = TRUE) # merge with full original dictionary to retrieve descriptions
icd_codes_to_be_double_checked_for_inclusion_gold = icd_codes_to_be_double_checked_for_inclusion_gold[order(-count_na)]
icd_codes_to_be_double_checked_for_inclusion_gold

# exclude events that refer to codes that are not part of final set of codes
gold_diagnoses_HES_APC_ICD_dict_final = rbind(gold_diagnoses_HES_APC_ICD_dict[!is.na(disease)], gold_diagnoses_HES_APC_ICD_dict_na[!is.na(disease)], fill = TRUE)
nrow(gold_diagnoses_HES_APC_ICD_dict_final) 
nrow(gold_diagnoses_HES_APC_ICD_dict_final[is.na(eventdate)])
nrow(gold_diagnoses_HES_APC_ICD_dict_final[is.na(disease)])

#OPCS table - Merge with code dictionary
gold_diagnoses_HES_APC_OPCS_dict = merge(gold_diagnoses_HES_APC_OPCS, opcs_codes_table[,.(alt_opcs_code, opcs_code, unique_ID, description, disease, disease_category, disease_subcategory, disease_short, include)], by.x = "opcs", by.y = "alt_opcs_code", all.x=T, allow.cartesian=TRUE) # ! no points in OPCS codes!
setnames(gold_diagnoses_HES_APC_OPCS_dict, old=c("opcs", "opcs_code", "unique_ID"), new =c("alt_code", "diag_code", "diag_code_unique_ID" )) # rename columns
nrow(gold_diagnoses_HES_APC_OPCS)
nrow(gold_diagnoses_HES_APC_OPCS_dict)
nrow(gold_diagnoses_HES_APC_OPCS_dict[is.na(disease)])
levels(as.factor(gold_diagnoses_HES_APC_OPCS_dict[is.na(disease)]$diag_code)) # Codes checked (these are codes that have been excluded from final analyses) 
gold_diagnoses_HES_APC_OPCS_dict = gold_diagnoses_HES_APC_OPCS_dict[!is.na(disease)]
nrow(gold_diagnoses_HES_APC_OPCS_dict)

# RBind ICD and OPCS tables
gold_diagnoses_HES_APC_dict = rbind(gold_diagnoses_HES_APC_ICD_dict_final,gold_diagnoses_HES_APC_OPCS_dict, fill=TRUE )
format(nrow(gold_diagnoses_HES_APC_dict), big.mark="'") 
```

### Diagnoses in CPRD GOLD - HES OP
```{r, eval=TRUE, warning=FALSE, message=FALSE}
# HES OP ICD
gold_diagnoses_HES_OP_ICD= fread(file.path(PATH_D,"/CPRD_GOLD/Event_of_interest/GOLD_HESOP_ICD_interim.txt.gz"), header = T, sep="\t")
gold_diagnoses_HES_OP_ICD_E05= fread(file.path(PATH_D,"/Extra_E_Chapter/Event_of_interest/GOLD_HESOP_ICD_interim_extra.txt.gz"), header = T, sep="\t") # additional E05 events
gold_diagnoses_HES_OP_ICD = rbind(gold_diagnoses_HES_OP_ICD, gold_diagnoses_HES_OP_ICD_E05)

nrow(gold_diagnoses_HES_OP_ICD)
gold_diagnoses_HES_OP_ICD[,database := "HES OP GOLD"] 
setnames(gold_diagnoses_HES_OP_ICD, old=c("apptdate"), new =c("eventdate")) # date of diagnosis = date of appointment 
gold_diagnoses_HES_OP_ICD[,eventdate:=as.Date(eventdate, format="%d/%m/%Y")] #set date format
nrow(gold_diagnoses_HES_OP_ICD[is.na(eventdate),]) # double check number of corrupt eventdates
gold_diagnoses_HES_OP_ICD = gold_diagnoses_HES_OP_ICD[!is.na(eventdate),] # Exclude corrupt eventdates
nrow(gold_diagnoses_HES_OP_ICD)
#nrow(gold_diagnoses_HES_OP_ICD[rapportools::is.empty(eventdate),]) # double check number of empty eventdates
gold_diagnoses_HES_OP_ICD = gold_diagnoses_HES_OP_ICD[duplicated(gold_diagnoses_HES_OP_ICD)==F,] # Exclude duplicates
nrow(gold_diagnoses_HES_OP_ICD)

gold_diagnoses_HES_OP_ICD[,codetype := "icd"] 
head(gold_diagnoses_HES_OP_ICD)

# HES OP OPCS
gold_diagnoses_HES_OP_OPCS= fread(file.path(PATH_D,"/CPRD_GOLD/Event_of_interest/GOLD_HESOP_OPCS_interim.txt.gz"), header = T, sep="\t")
nrow(gold_diagnoses_HES_OP_OPCS)
gold_diagnoses_HES_OP_OPCS[,database := "HES OP GOLD"] 
setnames(gold_diagnoses_HES_OP_OPCS, old=c("apptdate"), new =c("eventdate")) # date of diagnosis = date of appointment 
gold_diagnoses_HES_OP_OPCS[,eventdate:=as.Date(eventdate, format="%d/%m/%Y")] #set date format
nrow(gold_diagnoses_HES_OP_OPCS[is.na(eventdate),]) # double check number of corrupt eventdates
gold_diagnoses_HES_OP_OPCS = gold_diagnoses_HES_OP_OPCS[!is.na(eventdate),] # Exclude corrupt eventdates
nrow(gold_diagnoses_HES_OP_OPCS)
#nrow(gold_diagnoses_HES_OP_OPCS[rapportools::is.empty(eventdate),]) # double check number of empty eventdates
gold_diagnoses_HES_OP_OPCS = gold_diagnoses_HES_OP_OPCS[duplicated(gold_diagnoses_HES_OP_OPCS)==F,] # Exclude duplicates
nrow(gold_diagnoses_HES_OP_OPCS)
gold_diagnoses_HES_OP_OPCS[,codetype := "opcs"] 

#ICD table - Merge with code dictionary
gold_diagnoses_HES_OP_ICD_dict = merge(gold_diagnoses_HES_OP_ICD[,.(patid, eventdate, ICD, database, codetype)], icd_codes_table[,.(icd_code, alt_icd_code, unique_ID, description, disease, disease_category, disease_subcategory, disease_short, include)], by.x = "ICD", by.y = "alt_icd_code", all.x=T, allow.cartesian=TRUE)  # ! no points in OP codes!
head(gold_diagnoses_HES_OP_ICD_dict)
setnames(gold_diagnoses_HES_OP_ICD_dict, old=c("icd_code",  "ICD", "unique_ID"), new =c("diag_code","alt_code", "diag_code_unique_ID")) # rename columns
nrow(gold_diagnoses_HES_OP_ICD)
nrow(gold_diagnoses_HES_OP_ICD_dict)
nrow(gold_diagnoses_HES_OP_ICD_dict[is.na(disease)])
levels(as.factor(gold_diagnoses_HES_OP_ICD_dict[is.na(disease)]$diag_code)) # Codes checked (these are codes that have been excluded from final analyses)
gold_diagnoses_HES_OP_ICD_dict = gold_diagnoses_HES_OP_ICD_dict[!is.na(disease)]
nrow(gold_diagnoses_HES_OP_ICD_dict)

#OPCS table - Merge with code dictionary
gold_diagnoses_HES_OP_OPCS_dict = merge(gold_diagnoses_HES_OP_OPCS[,.(patid, eventdate, OPCS, database, codetype)], opcs_codes_table[,.(alt_opcs_code, opcs_code, unique_ID, description, disease, disease_category, disease_subcategory, disease_short, include)], by.x = "OPCS", by.y = "alt_opcs_code", all.x=T, allow.cartesian=TRUE) # ! no points in OPCS codes!
head(gold_diagnoses_HES_OP_OPCS_dict)
setnames(gold_diagnoses_HES_OP_OPCS_dict, old=c("OPCS", "opcs_code", "unique_ID"), new =c("alt_code", "diag_code", "diag_code_unique_ID")) # rename columns
nrow(gold_diagnoses_HES_OP_OPCS)
nrow(gold_diagnoses_HES_OP_OPCS_dict)
nrow(gold_diagnoses_HES_OP_OPCS_dict[is.na(disease)])
levels(as.factor(gold_diagnoses_HES_OP_OPCS_dict[is.na(disease)]$diag_code)) # Codes checked (these are codes that have been excluded from final analyses)

gold_diagnoses_HES_OP_OPCS_dict = gold_diagnoses_HES_OP_OPCS_dict[!is.na(disease)]
nrow(gold_diagnoses_HES_OP_OPCS_dict)

# RBind ICD and OPCS tables
gold_diagnoses_HES_OP_dict = rbind(gold_diagnoses_HES_OP_ICD_dict,gold_diagnoses_HES_OP_OPCS_dict, fill=TRUE )
nrow(gold_diagnoses_HES_OP_dict)


# !! NEW on 15 Sep 2021 - NEW GOLD patids
gold_diagnoses_HES_OP_dict[,patidold:= patid]
# change pracids
gold_diagnoses_HES_OP_dict[,pracid := as.character(substr(patid, floor(log10(patid))-1, floor(log10(patid))+1 ))] # pracid = last three digits of patidold
gold_diagnoses_HES_OP_dict[,pracid:= paste("10",pracid,sep="")] # new pracid = "10" + last three digits of patidold 
# change patids
gold_diagnoses_HES_OP_dict[,patid := as.character(substr(patidold, 1, floor(log10(patidold))-2 ))] # patid all digits up to the last three (which are the pracid)
gold_diagnoses_HES_OP_dict[,patid:= as.numeric(paste(patid,pracid,sep=""))] # new patid = old patid (up to the last three digits) + new pracid 
head(gold_diagnoses_HES_OP_dict)
```

### Diagnoses in CPRD GOLD - ONS Death 
```{r, eval=TRUE, warning=FALSE, message=FALSE}
death_gold= fread(file.path(PATH_D2,"/Gold/Data/Linked_Data/ONS/death_patient_19_156A_DM.txt.gz"), header = T, sep="\t")
format(nrow(death_gold), big.mark="'")  # number of records in table
class(death_gold$patid)
colnames(death_gold)
death_gold[,dod:=as.Date(dod, format="%d/%m/%Y")] #set date format
death_gold[,dod_partial:=as.Date(dod_partial, format="%d/%m/%Y")] #set date format
nrow(death_gold[is.na(dod)]) # check how many dod are missing
nrow(death_gold[is.na(dod) & !is.na(dod_partial)]) # check dod_partial can be used for those with missing dod
death_gold = death_gold[!is.na(dod)] # exclude records with missing death date
setnames(death_gold, old=c("cause", "cause1", "cause2", "cause3", "cause4", "cause5", "cause6", "cause7", "cause8", "cause9", "cause10", "cause11", "cause12", "cause13", "cause14", "cause15"), new =c("death_cause_first", "death_cause2", "death_cause3", "death_cause4", "death_cause5", "death_cause6", "death_cause7", "death_cause8", "death_cause9", "death_cause10", "death_cause11", "death_cause12", "death_cause13", "death_cause14", "death_cause15", "death_cause16"))
death_gold = death_gold[,.(patid, dod, death_cause_first, death_cause2, death_cause3, death_cause4, death_cause5, death_cause6, death_cause7, death_cause8, death_cause9, death_cause10, death_cause11, death_cause12, death_cause13, death_cause14, death_cause15, death_cause16)] # select relevant columns
head(death_gold)
format(nrow(death_gold), big.mark="'")  # number of records in table
format(length(unique(death_gold$patid)), big.mark="'")  # number of unique patient IDs in table

death_gold_melted = data.table::melt(death_gold, measure.vars = c("death_cause_first", "death_cause2", "death_cause3", "death_cause4", "death_cause5", "death_cause6", "death_cause7", "death_cause8", "death_cause9", "death_cause10", "death_cause11", "death_cause12", "death_cause13", "death_cause14", "death_cause15", "death_cause16"),variable.name = "death_diag_order", value.name = "diag_code") # melt table
#setnames(death_gold_melted, old = "value", new = "diag_code")
head(death_gold_melted)
death_gold_melted = death_gold_melted[!is.na(diag_code)] # exclude empty fields
death_gold_melted = death_gold_melted[!(diag_code == "")] # exclude empty fields
head(death_gold_melted)

format(nrow(death_gold_melted), big.mark="'")  # number of records in table
format(length(unique(death_gold_melted$patid)), big.mark="'")  # number of unique patient IDs in table

# Select only CVD codes
death_gold_melted_cvd = death_gold_melted[diag_code %in% icd_9_10_codes_table[disease_category == "cardiovascular"]$icd_code]
format(nrow(death_gold_melted_cvd), big.mark="'")  # number of records in table
format(length(unique(death_gold_melted_cvd$patid)), big.mark="'")  # number of unique patient IDs in table

# Merge with code dictionary
death_gold_melted_cvd2 = merge(death_gold_melted_cvd, icd_9_10_codes_table[disease_category == "cardiovascular",.(alt_icd_code, icd_code, unique_ID, description, disease, disease_category,disease_subcategory, disease_short, include)], by.x = "diag_code", by.y = "icd_code", all.x=T, allow.cartesian=TRUE)
levels(as.factor(fsetdiff(death_gold_melted_cvd[,.(diag_code)], death_gold_melted_cvd2[,.(diag_code)], all = TRUE)))
levels(as.factor(fsetdiff(death_gold_melted_cvd2[,.(diag_code)], death_gold_melted_cvd[,.(diag_code)], all = TRUE)))
icd_9_10_codes_table[icd_code %in% c("I26.0", "I26.9", "I81", "I82.3")] # these codes refer to 2 or more CVD
death_gold_melted_cvd = death_gold_melted_cvd2
setnames(death_gold_melted_cvd, old=c("alt_icd_code", "unique_ID"), new =c("alt_code", "diag_code_unique_ID")) # rename columns

# Check that all codes have been matched with a disease
nrow(death_gold_melted_cvd[is.na(disease)])

# Define code type and database
death_gold_melted_cvd[,database := "ONS GOLD"] 
death_gold_melted_cvd[,codetype := "icd"] 
levels(as.factor(death_gold_melted_cvd$death_diag_order))

# Rename dod as eventdate 
setnames(death_gold_melted_cvd, old=c("dod"), new =c("eventdate")) # rename columns
```

### Diagnoses in CPRD GOLD - HES PRIMARY (for sensitivity analyses)
```{r, eval=TRUE, warning=FALSE, message=FALSE}
hes_gold_primary= fread(file.path(PATH_D2,"/Gold/Data/Linked_Data/HES_Primary/hes_primary_diag_hosp_19_156A_DM.txt.gz"), header = T, sep="\t")
class(hes_gold_primary$patid)
head(hes_gold_primary)
hes_gold_primary[,admidate:=as.Date(admidate, format="%d/%m/%Y")] #set date format
hes_gold_primary[,discharged:=as.Date(discharged, format="%d/%m/%Y")] #set date format
format(nrow(hes_gold_primary), big.mark="'")  # number of records in table
format(nrow(hes_gold_primary[icdx != ""]), big.mark="'")  
hes_gold_primary[icdx %in% c("-","--","~","-~", "A", "D", "d", "Z", "-A","-D", "X", "E", "+", "*", ",", "!", "G", ".", "H", "F", "J", "K", "R", "N", "DN", "C", "I", "DH", "DR", "H3", "+L", "DG", "a", "M", "#", "~Y", "S", "0D", "2D", "2~", "9~","9D", " ", "0~")]$icdx = ""  # Exclude unhelpful modifiers
hes_gold_primary[icd_primary %in% c("I70.1", "I70.8", "I70.9",  "I48.0", "I48.1", "I48.9") & icdx %in% c("0", "1", "9")]$icdx = ""  # Exclude unhelpful modifiers

format(nrow(hes_gold_primary[icdx != ""]), big.mark="'") 
hes_gold_primary[,ICD_5:= paste(icd_primary, icdx, sep = "")] # concatenate ICD codes and modifiers as one single code
#icd_codes_table[disease_category == "cardiovascular"]
hes_gold_primary1 = subset(hes_gold_primary, icd_primary %in% icd_codes_table[disease_category == "cardiovascular"]$icd_code)
hes_gold_primary2 = subset(hes_gold_primary, ICD_5 %in% icd_codes_table[disease_category == "cardiovascular"]$icd_code)
nrow(hes_gold_primary1)
nrow(hes_gold_primary2)
nrow(hes_gold_primary1)-nrow(hes_gold_primary2)
fsetdiff(hes_gold_primary1, hes_gold_primary2, all = TRUE)
fsetdiff(hes_gold_primary2, hes_gold_primary1, all = TRUE)
hes_gold_primary = hes_gold_primary1
nrow(hes_gold_primary)

#comorb_gold_hes[,database := "HES APC gold"] 
#comorb_gold_hes[,codetype := "ICD"] 
setnames(hes_gold_primary, old=c("icd_primary", "icdx"), new =c("diag_code", "icd_modifier")) 
nrow(hes_gold_primary[is.na(discharged),]) # double check number of corrupt eventdates
hes_gold_primary[, eventdate:= discharged ] # date of diagnosis = date of discharge 
hes_gold_primary[is.na(eventdate), eventdate:= admidate ] # where discharge date is missing, use admission date
nrow(hes_gold_primary[is.na(eventdate),]) # double check number of corrupt eventdates
hes_gold_primary = hes_gold_primary[!is.na(eventdate)] # exclude events where both admission and discharge dates are missing
nrow(hes_gold_primary[is.na(eventdate)])

hes_gold_primary = hes_gold_primary[duplicated(hes_gold_primary)==F,] # Exclude duplicates
format(nrow(hes_gold_primary), big.mark="'") 
#comorb_gold_hes[,diag_code_unique_ID := paste("I",diag_code, sep="_")] ## Define unique_ID
colnames(hes_gold_primary)
hes_gold_primary = hes_gold_primary[,.(patid, eventdate, diag_code, admidate, discharged, spno, ICD_5, icd_modifier)]

# Merge with code dictionary
hes_gold_primary2 = merge(hes_gold_primary, icd_codes_table[disease_category == "cardiovascular",.(alt_icd_code, icd_code, unique_ID, description, disease, disease_category,disease_subcategory, disease_short, include)], by.x = "diag_code", by.y = "icd_code", all.x=T, allow.cartesian=TRUE)
levels(as.factor(fsetdiff(hes_gold_primary[,.(diag_code)], hes_gold_primary2[,.(diag_code)], all = TRUE)))
levels(as.factor(fsetdiff(hes_gold_primary2[,.(diag_code)], hes_gold_primary[,.(diag_code)], all = TRUE)))
icd_codes_table[icd_code %in% c("I26.0", "I26.9", "I81", "I82.3")]
hes_gold_primary = hes_gold_primary2
setnames(hes_gold_primary, old=c("alt_icd_code", "unique_ID"), new =c("alt_code", "diag_code_unique_ID")) # rename columns
colnames(hes_gold_primary)

# Check that all codes have been matched with a disease
nrow(hes_gold_primary[is.na(disease)])

# Define code type and database
hes_gold_primary[,database := "HES GOLD PRIMARY"] 
hes_gold_primary[,codetype := "icd"] 

# Add pracid
format(nrow(hes_gold_primary), big.mark="'")  # number of records in table
hes_gold_primary[,patid:=as.integer64(patid)]
hes_gold_primary = merge(hes_gold_primary, denominator_gold[,.(patid, pracid)], by = "patid", all.x = TRUE)
nrow(hes_gold_primary[is.na(pracid)]) # check that no pracids are missing

colnames(hes_gold_primary)
format(nrow(hes_gold_primary), big.mark="'")  # number of records in table
format(length(unique(hes_gold_primary$patid)), big.mark="'")  # number of unique patient IDs in table
```

### Combine all diagnoses in CPRD GOLD - Primary care, HES APC and HES OP, 
```{r combine_gold, eval=TRUE, warning=FALSE, message=FALSE}
include_cvd_death
if (include_cvd_death == "yes") {gold_diagnoses_all = rbind( gold_diagnoses_primary_care_dict, gold_diagnoses_HES_APC_dict,gold_diagnoses_HES_OP_dict, death_gold_melted_cvd, fill=TRUE ) }  else { gold_diagnoses_all = rbind( gold_diagnoses_primary_care_dict, gold_diagnoses_HES_APC_dict,gold_diagnoses_HES_OP_dict, fill=TRUE )} # if include_cvd_death == "yes" then include records of CVD death, else do not includes records of CVD death

format(nrow(gold_diagnoses_all), big.mark="'") 
head(gold_diagnoses_all)

# Check that for all diag_codes include == 1
levels(as.factor(gold_diagnoses_all$include))

# check patids
summary(gold_diagnoses_all$patid)
class(gold_diagnoses_all$patid)
nrow(gold_diagnoses_all[patid == 1002210231]) # new patid, should be in the table
nrow(gold_diagnoses_all[patid == 10231]) # old patid, should not appear in the table
nrow(gold_diagnoses_all[patid == 1000010010]) # new patid, should be in the table
nrow(gold_diagnoses_all[patid == 10010]) # old patid, should not appear in the table

summary(denominator_gold$patid)
class(denominator_gold$patid)
denominator_gold[,patid := as.integer64(patid)]
head(denominator_gold)
nrow(denominator_gold[patid == 1002210231]) # new patid, should be in the table
nrow(denominator_gold[patid == 10231]) # old patid, should not appear in the table
nrow(denominator_gold[patid == 1000010010]) # new patid, should be in the table
nrow(denominator_gold[patid == 10010]) # old patid, should not appear in the table
#1000010010	10010	10000010	10	
#1002210231	10231	10022231	231


# Deduplicate patids that are in both aurum and gold (the are marked with exclude ==1 in gold data)
gold_diagnoses_all = merge(gold_diagnoses_all, denominator_gold[,.(patid, exclude)], by = "patid", all.x = TRUE)
nrow(gold_diagnoses_all)
nrow(denominator_gold[is.na(exclude)])
nrow(denominator_gold[exclude == 1])
nrow(gold_diagnoses_all) - nrow(gold_diagnoses_all[exclude == 1]) -nrow(gold_diagnoses_all[exclude == 0]) # check that this is 0
gold_diagnoses_all = gold_diagnoses_all[exclude == 0] ## exclude patids with exclude == 1 - those patients who are both in CPRD gold and aurum
format(nrow(gold_diagnoses_all), big.mark="'") 

# Check eventdates and patid
nrow(gold_diagnoses_all[is.na(eventdate),])
nrow(gold_diagnoses_all[is.na(pracid),])

# define db_patid across gold and aurum
gold_diagnoses_all[,db_patid := paste("G", patid, sep="")] 
```

## Save_0
```{r save_0, eval=TRUE, warning=FALSE, message=FALSE}
#save.image(file.path(PATH_W,"2022-02-18SelectedWorkspaceIdentification_0.RData"))
rm(list=setdiff(ls(), c("gold_diagnoses_all",  "aurum_diagnoses_all", "death_gold_melted_cvd", "death_aurum_melted_cvd", "hes_gold_primary", "hes_aurum_primary", "study_start_date", "study_end_date", "PATH", "PATH_D", "PATH_D2", "PATH_W", "PATH_W_orig", "PATH_L", "gold_codes_table", "aurum_codes_table", "icd_codes_table", "icd_9_10_codes_table", "opcs_codes_table", "all_codes_table", "denominator", "denominator_aurum", "age_group_table", "compute_days", "age_group_fun", "icd_codes_to_be_double_checked_for_inclusion_gold", "diag_codes_table_inc_0", "gold_codes_to_be_double_checked_for_inclusion", "aurum_codes_to_be_double_checked_for_inclusion", "icd_codes_to_be_double_checked_for_inclusion_aurum", "include_cvd_death", "drug_codes", "drug_codes_gold", "drug_codes_aurum", "death_gold", "death_aurum")))
#load(file.path(PATH_W,"2022-02-18SelectedWorkspaceIdentification_0.RData"))
#save.image(file.path(PATH_W,"2022-02-18SelectedWorkspaceIdentification_0.RData"))
gc()
```

## 2.2 Diagnoses in CPRD Aurum
### Diagnoses in CPRD Aurum - Primary Care - !! Needs to be updated with the latest "CVD_INCIDENCE_AURUM_OBSERVATION_diagnoses.Rmd" file when diagnostic codes change !!
```{r, eval=TRUE, warning=FALSE, message=FALSE}
aurum_diagnoses_primary_care= fread(file.path(PATH_W,"/observation_aurum_diagnoses.txt.gz"), colClasses=c("patid"="integer64", "medcodeid"="integer64", "obsdate"="Date"))
format(nrow(aurum_diagnoses_primary_care), big.mark="'")  # Check total number of events
head(aurum_diagnoses_primary_care)
setnames(aurum_diagnoses_primary_care, old=c( "obsdate", "medcodeid"), new =c("eventdate", "medcode_aurum")) # rename columns
class(aurum_diagnoses_primary_care$medcode_aurum) # check class is integer64 (CPRD Aurum only)
#aurum_diagnoses_primary_care[,eventdate:=as.Date(eventdate, format="%d/%m/%Y")] # Set date format
format(nrow(aurum_diagnoses_primary_care[is.na(eventdate),]), big.mark="'") # Check number of corrupt eventdates
aurum_diagnoses_primary_care=aurum_diagnoses_primary_care[!is.na(eventdate),] # Exclude events with corrupt date
#nrow(aurum_diagnoses_primary_care[rapportools::is.empty(eventdate),]) # Check number of empty eventdates
format(nrow(aurum_diagnoses_primary_care), big.mark="'") 
aurum_diagnoses_primary_care = aurum_diagnoses_primary_care[duplicated(aurum_diagnoses_primary_care)==F,] # Exclude duplicates
format(nrow(aurum_diagnoses_primary_care), big.mark="'") 
aurum_diagnoses_primary_care[,database := "CPRD Aurum"]
aurum_diagnoses_primary_care[,codetype := "aurum"]
head(aurum_diagnoses_primary_care)

#Merge with code dictionary
aurum_diagnoses_primary_care_dict = merge(aurum_diagnoses_primary_care, aurum_codes_table[,!c("database","codetype", "keyword")], by = "medcode_aurum", all.x=T, allow.cartesian=TRUE)
setnames(aurum_diagnoses_primary_care_dict, old=c( "medcode_aurum", "unique_ID"), new =c("diag_code_aurum", "diag_code_unique_ID")) # rename to diag_code
head(aurum_diagnoses_primary_care_dict)
colnames(aurum_diagnoses_primary_care_dict)
nrow(aurum_diagnoses_primary_care)
nrow(aurum_diagnoses_primary_care_dict)
nrow(aurum_diagnoses_primary_care_dict[is.na(disease)])
levels(as.factor(aurum_diagnoses_primary_care_dict[is.na(disease)]$diag_code_aurum)) # Codes checked (these are codes that have been excluded from final analyses) 
levels(as.factor(aurum_diagnoses_primary_care_dict$disease)) # Codes checked (these are codes that have been excluded from final analyses) 

# check which codes remain invalid
aurum_codes_to_be_double_checked_for_inclusion = aurum_diagnoses_primary_care_dict[is.na(disease)]
aurum_codes_to_be_double_checked_for_inclusion[, count_na:= .N, by = c("diag_code_aurum")]
aurum_codes_to_be_double_checked_for_inclusion = unique(aurum_codes_to_be_double_checked_for_inclusion[,.(diag_code_aurum, count_na)])
aurum_codes_to_be_double_checked_for_inclusion = merge(aurum_codes_to_be_double_checked_for_inclusion, diag_codes_table_inc_0[codetype == "aurum",.(description, disease, disease_category, unique_ID, medcode_aurum,codetype, include)], by.x = "diag_code_aurum", by.y = "medcode_aurum", all.x = TRUE) # merge with full original dictionary to retrieve descriptions
aurum_codes_to_be_double_checked_for_inclusion = aurum_codes_to_be_double_checked_for_inclusion[order(-count_na)]
aurum_codes_to_be_double_checked_for_inclusion

aurum_diagnoses_primary_care_dict = aurum_diagnoses_primary_care_dict[!is.na(disease)]
nrow(aurum_diagnoses_primary_care_dict)
```

### Diagnoses in CPRD Aurum - HES APC
```{r aurum_apc, eval=TRUE, warning=FALSE, message=FALSE}
aurum_diagnoses_HES_APC_ICD= fread(file.path(PATH_D2,"/Aurum/Data/Linked_Data/HES_Diagnoses_by_hosp/hes_diagnosis_hosp_19_156A_DM.txt.gz"), header = T, sep="\t")
head(aurum_diagnoses_HES_APC_ICD)
nrow(aurum_diagnoses_HES_APC_ICD)
colnames(aurum_diagnoses_HES_APC_ICD)
aurum_diagnoses_HES_APC_ICD[as.numeric(ICD)>0] # double-check if icd-9 codes were used (icd-9 codes are numeric and should not have been used here)

aurum_diagnoses_HES_APC_ICD[,database := "HES APC Aurum"]
setnames(aurum_diagnoses_HES_APC_ICD, old=c("discharged", "ICD", "ICDx"), new =c("eventdate", "ICD_main", "ICD_modifier")) # date of diagnosis = date of admission 
aurum_diagnoses_HES_APC_ICD[,eventdate:=as.Date(eventdate, format="%d/%m/%Y")] #set date format
nrow(aurum_diagnoses_HES_APC_ICD[is.na(eventdate),]) # double check number of corrupt eventdates
aurum_diagnoses_HES_APC_ICD[is.na(eventdate), eventdate:= as.Date(admidate, format="%d/%m/%Y")] # if discharge date is missing, then use epiend date
#aurum_diagnoses_HES_APC_ICD[is.na(eventdate), eventdate:= as.Date(epistart, format="%d/%m/%Y")] # if discharge date is missing, then use epistart date
nrow(aurum_diagnoses_HES_APC_ICD[is.na(eventdate),]) # double check number of corrupt eventdates
#aurum_diagnoses_HES_APC_ICD[is.na(eventdate),]
aurum_diagnoses_HES_APC_ICD = aurum_diagnoses_HES_APC_ICD[!is.na(eventdate),] # Exclude corrupt eventdates
nrow(aurum_diagnoses_HES_APC_ICD)
#nrow(aurum_diagnoses_HES_APC_ICD[rapportools::is.empty(eventdate),]) # double check number of empty eventdates
aurum_diagnoses_HES_APC_ICD = aurum_diagnoses_HES_APC_ICD[duplicated(aurum_diagnoses_HES_APC_ICD)==F,] # Exclude duplicates
nrow(aurum_diagnoses_HES_APC_ICD)
aurum_diagnoses_HES_APC_ICD[,codetype := "icd"] 
aurum_diagnoses_HES_APC_ICD[ICD_modifier %in% c("-","--","~","-~", "A", "D", "d", "Z", "-A","-D", "X", "E", "+", "*", ",", "!", "G", ".", "H", "F", "J", "K", "R", "N", "DN", "C", "I", "DH", "DR", "H3", "+L", "DG", "a", "M", "#", "~Y", "S", "0D", "2D", "2~", "9~","9D")]$ICD_modifier = ""  # Exclude unhelpful modifiers
aurum_diagnoses_HES_APC_ICD$ICD = paste(aurum_diagnoses_HES_APC_ICD$ICD_main, aurum_diagnoses_HES_APC_ICD$ICD_modifier, sep="") # concatenate ICD codes and modifiers as one single code
aurum_diagnoses_HES_APC_ICD = unique(aurum_diagnoses_HES_APC_ICD[,.(patid, eventdate, ICD, database, codetype)]) 
nrow(aurum_diagnoses_HES_APC_ICD)
#nrow(aurum_diagnoses_HES_APC_ICD[is.na(eventdate)])
#nrow(aurum_diagnoses_HES_APC_ICD[is.na(pracid)])

# HES APC OPCS
aurum_diagnoses_HES_APC_OPCS= fread(file.path(PATH_D,"/CPRD_Aurum/Event_of_interest/Aurum_HESAPC_OPCS_interim.txt.gz"), header = T, sep="\t")
nrow(aurum_diagnoses_HES_APC_OPCS)
aurum_diagnoses_HES_APC_OPCS[,database := "HES APC Aurum"] 

setnames(aurum_diagnoses_HES_APC_OPCS, old=c("discharged"), new =c("eventdate")) # date of diagnosis = date of admission 
aurum_diagnoses_HES_APC_OPCS[,eventdate:=as.Date(eventdate, format="%d/%m/%Y")] #set date format
nrow(aurum_diagnoses_HES_APC_OPCS[is.na(eventdate),]) # double check number of corrupt eventdates
aurum_diagnoses_HES_APC_OPCS[is.na(eventdate), eventdate:= as.Date(epiend, format="%d/%m/%Y")] # if discharge date is missing, then use epiend date
aurum_diagnoses_HES_APC_OPCS[is.na(eventdate), eventdate:= as.Date(epistart, format="%d/%m/%Y")] # if discharge date is missing, then use epistart date
nrow(aurum_diagnoses_HES_APC_OPCS[is.na(eventdate),]) # double check number of corrupt eventdates
aurum_diagnoses_HES_APC_OPCS = aurum_diagnoses_HES_APC_OPCS[!is.na(eventdate),] # Exclude corrupt eventdates
nrow(aurum_diagnoses_HES_APC_OPCS)
#nrow(aurum_diagnoses_HES_APC_OPCS[rapportools::is.empty(eventdate),]) # double check number of empty eventdates
aurum_diagnoses_HES_APC_OPCS[,codetype := "opcs"] 
aurum_diagnoses_HES_APC_OPCS = unique(aurum_diagnoses_HES_APC_OPCS[,.(patid, eventdate, OPCS, database, codetype)]) # Exclude duplicates
nrow(aurum_diagnoses_HES_APC_OPCS)

#ICD table - Merge with code dictionary
aurum_diagnoses_HES_APC_ICD_dict = merge(aurum_diagnoses_HES_APC_ICD, icd_codes_table[,.(alt_icd_code, icd_code, unique_ID, description, disease, disease_category, disease_subcategory, disease_short, include)], by.x = "ICD", by.y = "icd_code", all=T, allow.cartesian=TRUE)
setnames(aurum_diagnoses_HES_APC_ICD_dict, old=c("ICD",  "alt_icd_code", "unique_ID"), new =c("diag_code","alt_code", "diag_code_unique_ID")) # rename columns
aurum_diagnoses_HES_APC_ICD_dict[,orig_diag_code := diag_code]
nrow(aurum_diagnoses_HES_APC_ICD)
nrow(aurum_diagnoses_HES_APC_ICD_dict)
nrow(aurum_diagnoses_HES_APC_ICD_dict[is.na(disease)])
levels(as.factor(aurum_diagnoses_HES_APC_ICD_dict[is.na(disease)]$diag_code)) # Codes checked (these are codes that have been excluded from final analyses) 

## CHECK
kable(aurum_diagnoses_HES_APC_ICD_dict[is.na(eventdate) & !is.na(disease)]) # ICD codes for which there have been no match!
aurum_diagnoses_HES_APC_ICD_dict = aurum_diagnoses_HES_APC_ICD_dict[!is.na(eventdate)] # exclude records missing eventdates (here these are ICD codes for which there have been no matches (ICD chapter codes))
nrow(aurum_diagnoses_HES_APC_ICD_dict[is.na(eventdate)]) 

# for codes that do not match, try to match with removing last character
aurum_diagnoses_HES_APC_ICD_dict_na = aurum_diagnoses_HES_APC_ICD_dict[is.na(disease)]
aurum_diagnoses_HES_APC_ICD_dict_na[,short_diag_code := substr(as.character(diag_code), 1, nchar(as.character(diag_code))-1)]
aurum_diagnoses_HES_APC_ICD_dict_na = merge(unique(aurum_diagnoses_HES_APC_ICD_dict_na[,.(patid, eventdate, diag_code, short_diag_code, database, codetype)]), icd_codes_table[,.(icd_code, unique_ID, description, disease, disease_category,disease_subcategory, disease_short, include)], by.x = "short_diag_code", by.y = "icd_code", all.x=T, allow.cartesian=TRUE)
head(aurum_diagnoses_HES_APC_ICD_dict_na[!is.na(disease)])
aurum_diagnoses_HES_APC_ICD_dict_na[,diag_code_orig:=diag_code]
aurum_diagnoses_HES_APC_ICD_dict_na[,diag_code:=short_diag_code]
aurum_diagnoses_HES_APC_ICD_dict_na[,diag_code_unique_ID:=unique_ID]
nrow(aurum_diagnoses_HES_APC_ICD_dict_na)
nrow(aurum_diagnoses_HES_APC_ICD_dict_na[is.na(disease)])
aurum_diagnoses_HES_APC_ICD_dict_na[is.na(disease)]

# check which codes remain invalid
icd_codes_to_be_double_checked_for_inclusion_aurum = aurum_diagnoses_HES_APC_ICD_dict_na[is.na(disease)]
icd_codes_to_be_double_checked_for_inclusion_aurum[, count_na:= .N, by = c("diag_code")]
icd_codes_to_be_double_checked_for_inclusion_aurum = unique(icd_codes_to_be_double_checked_for_inclusion_aurum[,.(diag_code, short_diag_code, count_na)])
icd_codes_to_be_double_checked_for_inclusion_aurum = merge(icd_codes_to_be_double_checked_for_inclusion_aurum, diag_codes_table_inc_0[codetype == "icd",.(description, disease, disease_category, unique_ID, icd_code,codetype, include)], by.x = "diag_code", by.y = "icd_code", all.x = TRUE) # merge with full original dictionary to retrieve descriptions
icd_codes_to_be_double_checked_for_inclusion_aurum = icd_codes_to_be_double_checked_for_inclusion_aurum[order(-count_na)]
icd_codes_to_be_double_checked_for_inclusion_aurum

# exclude events that refer to codes that are not part of final set of codes
aurum_diagnoses_HES_APC_ICD_dict_final = rbind(aurum_diagnoses_HES_APC_ICD_dict[!is.na(disease)], aurum_diagnoses_HES_APC_ICD_dict_na[!is.na(disease)], fill = TRUE)
nrow(aurum_diagnoses_HES_APC_ICD_dict_final) 
nrow(aurum_diagnoses_HES_APC_ICD_dict_final[is.na(eventdate)])
nrow(aurum_diagnoses_HES_APC_ICD_dict_final[is.na(disease)])

#OPCS table - Merge with code dictionary
aurum_diagnoses_HES_APC_OPCS_dict = merge(aurum_diagnoses_HES_APC_OPCS[,.(patid, eventdate, OPCS, database, codetype)], opcs_codes_table[,.(alt_opcs_code, opcs_code, unique_ID, description, disease, disease_category, disease_subcategory, disease_short, include)], by.x = "OPCS", by.y = "alt_opcs_code", all.x=T, allow.cartesian=TRUE) # ! no points in OPCS codes!
setnames(aurum_diagnoses_HES_APC_OPCS_dict, old=c("OPCS", "opcs_code", "unique_ID"), new =c("alt_code", "diag_code", "diag_code_unique_ID")) # rename columns
nrow(aurum_diagnoses_HES_APC_OPCS)
nrow(aurum_diagnoses_HES_APC_OPCS_dict)
nrow(aurum_diagnoses_HES_APC_OPCS_dict[is.na(disease)])
levels(as.factor(aurum_diagnoses_HES_APC_OPCS_dict[is.na(disease)]$diag_code)) # Codes checked (these are codes that have been excluded from final analyses) 
aurum_diagnoses_HES_APC_OPCS_dict = aurum_diagnoses_HES_APC_OPCS_dict[!is.na(disease)]
nrow(aurum_diagnoses_HES_APC_OPCS_dict)
nrow(aurum_diagnoses_HES_APC_OPCS_dict[is.na(eventdate)])
nrow(aurum_diagnoses_HES_APC_OPCS_dict[is.na(disease)])

# RBind ICD and OPCS tables
aurum_diagnoses_HES_APC_dict = rbind(aurum_diagnoses_HES_APC_ICD_dict_final,aurum_diagnoses_HES_APC_OPCS_dict, fill=TRUE )
```

### Diagnoses in CPRD Aurum - HES OP
```{r, eval=TRUE, warning=FALSE, message=FALSE}
# HES OP ICD
aurum_diagnoses_HES_OP_ICD= fread(file.path(PATH_D,"/CPRD_Aurum/Event_of_interest/Aurum_HESOP_ICD_interim.txt.gz"), header = T, sep="\t")
aurum_diagnoses_HES_OP_ICD_E05= fread(file.path(PATH_D,"/Extra_E_Chapter/Event_of_interest/Aurum_HESOP_ICD_interim_extra.txt.gz"), header = T, sep="\t") # additional E05 events
nrow(aurum_diagnoses_HES_OP_ICD_E05)
aurum_diagnoses_HES_OP_ICD = rbind(aurum_diagnoses_HES_OP_ICD, aurum_diagnoses_HES_OP_ICD_E05)

nrow(aurum_diagnoses_HES_OP_ICD)
aurum_diagnoses_HES_OP_ICD[,database := "HES OP Aurum"] 
setnames(aurum_diagnoses_HES_OP_ICD, old=c("apptdate"), new =c("eventdate")) # date of diagnosis = date of appointment 
aurum_diagnoses_HES_OP_ICD[,eventdate:=as.Date(eventdate, format="%d/%m/%Y")] #set date format
nrow(aurum_diagnoses_HES_OP_ICD[is.na(eventdate),]) # double check number of corrupt eventdates
aurum_diagnoses_HES_OP_ICD = aurum_diagnoses_HES_OP_ICD[!is.na(eventdate),] # Exclude corrupt eventdates
nrow(aurum_diagnoses_HES_OP_ICD)
#nrow(aurum_diagnoses_HES_OP_ICD[rapportools::is.empty(eventdate),]) # double check number of empty eventdates
#aurum_diagnoses_HES_OP_ICD = aurum_diagnoses_HES_OP_ICD[duplicated(aurum_diagnoses_HES_OP_ICD)==F,] # Exclude duplicates
nrow(aurum_diagnoses_HES_OP_ICD)
aurum_diagnoses_HES_OP_ICD[,codetype := "icd"] 
aurum_diagnoses_HES_OP_ICD=aurum_diagnoses_HES_OP_ICD[,.(patid, eventdate, ICD, database, codetype)] # Exclude duplicates
nrow(aurum_diagnoses_HES_OP_ICD)
head(aurum_diagnoses_HES_OP_ICD)

# HES OP OPCS
aurum_diagnoses_HES_OP_OPCS= fread(file.path(PATH_D,"/CPRD_Aurum/Event_of_interest/Aurum_HESOP_OPCS_interim.txt.gz"), header = T, sep="\t")
nrow(aurum_diagnoses_HES_OP_OPCS)
aurum_diagnoses_HES_OP_OPCS[,database := "HES OP Aurum"] 
setnames(aurum_diagnoses_HES_OP_OPCS, old=c("apptdate"), new =c("eventdate")) # date of diagnosis = date of appointment 
aurum_diagnoses_HES_OP_OPCS[,eventdate:=as.Date(eventdate, format="%d/%m/%Y")] #set date format
nrow(aurum_diagnoses_HES_OP_OPCS[is.na(eventdate),]) # double check number of corrupt eventdates
aurum_diagnoses_HES_OP_OPCS = aurum_diagnoses_HES_OP_OPCS[!is.na(eventdate),] # Exclude corrupt eventdates
nrow(aurum_diagnoses_HES_OP_OPCS)
#nrow(aurum_diagnoses_HES_OP_OPCS[rapportools::is.empty(eventdate),]) # double check number of empty eventdates
#aurum_diagnoses_HES_OP_OPCS = aurum_diagnoses_HES_OP_OPCS[duplicated(aurum_diagnoses_HES_OP_OPCS)==F,] # Exclude duplicates
nrow(aurum_diagnoses_HES_OP_OPCS)
aurum_diagnoses_HES_OP_OPCS[,codetype := "opcs"] 
aurum_diagnoses_HES_OP_OPCS=aurum_diagnoses_HES_OP_OPCS[,.(patid, eventdate, OPCS, database, codetype)] # Exclude duplicates
nrow(aurum_diagnoses_HES_OP_OPCS)
head(aurum_diagnoses_HES_OP_OPCS)

#ICD table - Merge with code dictionary
aurum_diagnoses_HES_OP_ICD_dict = merge(aurum_diagnoses_HES_OP_ICD[,.(patid, eventdate, ICD, database, codetype)], icd_codes_table[,.(alt_icd_code, icd_code, unique_ID, description, disease, disease_category, disease_subcategory, disease_short, include)], by.x = "ICD", by.y = "alt_icd_code", all.x=T, allow.cartesian=TRUE)  # ! no points in OP codes!
head(aurum_diagnoses_HES_OP_ICD_dict)
setnames(aurum_diagnoses_HES_OP_ICD_dict, old=c("icd_code",  "ICD", "unique_ID"), new =c("diag_code","alt_code", "diag_code_unique_ID")) # rename columns
nrow(aurum_diagnoses_HES_OP_ICD)
nrow(aurum_diagnoses_HES_OP_ICD_dict)
nrow(aurum_diagnoses_HES_OP_ICD_dict[is.na(disease)])
aurum_diagnoses_HES_OP_ICD_dict[is.na(disease)]
levels(as.factor(aurum_diagnoses_HES_OP_ICD_dict[is.na(disease)]$diag_code)) # Codes checked (these are codes that have been excluded from final analyses) 
aurum_diagnoses_HES_OP_ICD_dict = aurum_diagnoses_HES_OP_ICD_dict[!is.na(disease)]
nrow(aurum_diagnoses_HES_OP_ICD_dict)

#OPCS table - Merge with code dictionary
aurum_diagnoses_HES_OP_OPCS_dict = merge(aurum_diagnoses_HES_OP_OPCS[,.(patid, eventdate, OPCS, database, codetype)], opcs_codes_table[,.(alt_opcs_code, opcs_code, unique_ID, description, disease, disease_category, disease_subcategory, disease_short, include)], by.x = "OPCS", by.y = "alt_opcs_code", all.x=T, allow.cartesian=TRUE) # ! no points in OPCS codes!
head(aurum_diagnoses_HES_OP_OPCS_dict)
setnames(aurum_diagnoses_HES_OP_OPCS_dict, old=c("OPCS", "opcs_code", "unique_ID"), new =c("alt_code", "diag_code", "diag_code_unique_ID")) # rename columns
nrow(aurum_diagnoses_HES_OP_OPCS)
nrow(aurum_diagnoses_HES_OP_OPCS_dict)
nrow(aurum_diagnoses_HES_OP_OPCS_dict[is.na(disease)])
levels(as.factor(aurum_diagnoses_HES_OP_OPCS_dict[is.na(disease)]$diag_code)) # Codes checked (these are codes that have been excluded from final analyses) 
aurum_diagnoses_HES_OP_OPCS_dict = aurum_diagnoses_HES_OP_OPCS_dict[!is.na(disease)]
nrow(aurum_diagnoses_HES_OP_OPCS_dict)

# RBind ICD and OPCS tables
aurum_diagnoses_HES_OP_dict = rbind(aurum_diagnoses_HES_OP_ICD_dict,aurum_diagnoses_HES_OP_OPCS_dict, fill=TRUE )
#head(aurum_diagnoses_HES_OP_OPCS_dict)
nrow(aurum_diagnoses_HES_OP_dict)
nrow(aurum_diagnoses_HES_OP_OPCS_dict)+nrow(aurum_diagnoses_HES_OP_ICD_dict)
```

### Diagnoses in CPRD Aurum - ONS Death
```{r, eval=TRUE, warning=FALSE, message=FALSE}
death_aurum= fread(file.path(PATH_D2,"/Aurum/Data/Linked_Data/ONS/death_patient_19_156A_DM.txt.gz"), header = T, sep="\t")
class(death_aurum$patid)
colnames(death_aurum)
death_aurum[,dod:=as.Date(dod, format="%d/%m/%Y")] #set date format
death_aurum[,dod_partial:=as.Date(dod_partial, format="%d/%m/%Y")] #set date format
nrow(death_aurum[is.na(dod)]) # check how many dod are missing
nrow(death_aurum[is.na(dod) & !is.na(dod_partial)]) # check dod_partial can be used for those with missing dod
death_aurum = death_aurum[!is.na(dod)] # exclude records with missing death date
setnames(death_aurum, old=c("cause", "cause1", "cause2", "cause3", "cause4", "cause5", "cause6", "cause7", "cause8", "cause9", "cause10", "cause11", "cause12", "cause13", "cause14", "cause15"), new =c("death_cause_first", "death_cause2", "death_cause3", "death_cause4", "death_cause5", "death_cause6", "death_cause7", "death_cause8", "death_cause9", "death_cause10", "death_cause11", "death_cause12", "death_cause13", "death_cause14", "death_cause15", "death_cause16"))
death_aurum = death_aurum[,.(patid, dod, death_cause_first, death_cause2, death_cause3, death_cause4, death_cause5, death_cause6, death_cause7, death_cause8, death_cause9, death_cause10, death_cause11, death_cause12, death_cause13, death_cause14, death_cause15, death_cause16)] # select relevant columns
head(death_aurum)
format(nrow(death_aurum), big.mark="'")  # number of records in table
format(length(unique(death_aurum$patid)), big.mark="'")  # number of unique patient IDs in table

death_aurum_melted = data.table::melt(death_aurum, measure.vars = c("death_cause_first", "death_cause2", "death_cause3", "death_cause4", "death_cause5", "death_cause6", "death_cause7", "death_cause8", "death_cause9", "death_cause10", "death_cause11", "death_cause12", "death_cause13", "death_cause14", "death_cause15", "death_cause16"),variable.name = "death_diag_order", value.name = "diag_code") # melt table
#setnames(death_aurum_melted, old = "value", new = "diag_code")
death_aurum_melted = death_aurum_melted[!is.na(diag_code)] # exclude empty fields
death_aurum_melted = death_aurum_melted[!(diag_code == "")] # exclude empty fields
head(death_aurum_melted)

format(nrow(death_aurum_melted), big.mark="'")  # number of records in table
format(length(unique(death_aurum_melted$patid)), big.mark="'")  # number of unique patient IDs in table

# Select only CVD codes
death_aurum_melted_cvd = death_aurum_melted[diag_code %in% icd_9_10_codes_table[disease_category == "cardiovascular"]$icd_code]
format(nrow(death_aurum_melted_cvd), big.mark="'")  # number of records in table
format(length(unique(death_aurum_melted_cvd$patid)), big.mark="'")  # number of unique patient IDs in table

# Merge with code dictionary
death_aurum_melted_cvd2 = merge(death_aurum_melted_cvd, icd_9_10_codes_table[disease_category == "cardiovascular",.(alt_icd_code, icd_code, unique_ID, description, disease, disease_category,disease_subcategory, disease_short, include)], by.x = "diag_code", by.y = "icd_code", all.x=T, allow.cartesian=TRUE)
levels(as.factor(fsetdiff(death_aurum_melted_cvd[,.(diag_code)], death_aurum_melted_cvd2[,.(diag_code)], all = TRUE)))
levels(as.factor(fsetdiff(death_aurum_melted_cvd2[,.(diag_code)], death_aurum_melted_cvd[,.(diag_code)], all = TRUE)))
icd_9_10_codes_table[icd_code %in% c("I26.0", "I26.9", "I81", "I82.3")]
death_aurum_melted_cvd = death_aurum_melted_cvd2
setnames(death_aurum_melted_cvd, old=c("alt_icd_code", "unique_ID"), new =c("alt_code", "diag_code_unique_ID")) # rename columns

# Check that all codes have been matched with a disease
nrow(death_aurum_melted_cvd[is.na(disease)])

# Define code type and database
death_aurum_melted_cvd[,database := "ONS AURUM"] 
death_aurum_melted_cvd[,codetype := "icd"] 

# Rename dod as eventdate 
setnames(death_aurum_melted_cvd, old=c("dod"), new =c("eventdate")) # rename columns

format(nrow(death_aurum_melted_cvd), big.mark="'")  # number of records in table
colnames(death_aurum_melted_cvd)
```

### Diagnoses in CPRD Aurum - HES PRIMARY (for sensitivity analyses)
```{r, eval=TRUE, warning=FALSE, message=FALSE}
hes_aurum_primary= fread(file.path(PATH_D2,"/Aurum/Data/Linked_Data/HES_Primary/hes_primary_diag_hosp_19_156A_DM.txt.gz"), header = T, sep="\t")
class(hes_aurum_primary$patid)
head(hes_aurum_primary)
hes_aurum_primary[,admidate:=as.Date(admidate, format="%d/%m/%Y")] #set date format
hes_aurum_primary[,discharged:=as.Date(discharged, format="%d/%m/%Y")] #set date format
format(nrow(hes_aurum_primary), big.mark="'")  # number of records in table
format(nrow(hes_aurum_primary[ICDx != ""]), big.mark="'")  
hes_aurum_primary[ICDx %in% c("-","--","~","-~", "A", "D", "d", "Z", "-A","-D", "X", "E", "+", "*", ",", "!", "G", ".", "H", "F", "J", "K", "R", "N", "DN", "C", "I", "DH", "DR", "H3", "+L", "DG", "a", "M", "#", "~Y", "S", "0D", "2D", "2~", "9~","9D", " ", "0~")]$ICDx = ""  # Exclude unhelpful modifiers
hes_aurum_primary[ICD_PRIMARY %in% c("I70.1", "I70.8", "I70.9",  "I48.0", "I48.1", "I48.9") & ICDx %in% c("0", "1", "9")]$ICDx = ""  # Exclude unhelpful modifiers
#nrow(hes_aurum_primary[ICD_PRIMARY == "I70.1"])
#nrow(hes_aurum_primary[ICD_PRIMARY == "I70.10"])
#nrow(hes_aurum_primary[ICD_PRIMARY == "I70.9"])
#nrow(hes_aurum_primary[ICD_PRIMARY == "I70.8"])
#nrow(hes_aurum_primary[ICD_PRIMARY == "I70.80"])
#nrow(hes_aurum_primary[ICD_PRIMARY == "I48.9"])
#nrow(hes_aurum_primary[ICD_PRIMARY == "I48.90"])

format(nrow(hes_aurum_primary[ICDx != ""]), big.mark="'") 
hes_aurum_primary[,ICD_5:= paste(ICD_PRIMARY, ICDx, sep = "")] # concatenate ICD codes and modifiers as one single code
#icd_codes_table[disease_category == "cardiovascular"]
hes_aurum_primary1 = subset(hes_aurum_primary, ICD_PRIMARY %in% icd_codes_table[disease_category == "cardiovascular"]$icd_code)
hes_aurum_primary2 = subset(hes_aurum_primary, ICD_5 %in% icd_codes_table[disease_category == "cardiovascular"]$icd_code)
nrow(hes_aurum_primary1)
nrow(hes_aurum_primary2)
nrow(hes_aurum_primary1)-nrow(hes_aurum_primary2)
fsetdiff(hes_aurum_primary1, hes_aurum_primary2, all = TRUE)
fsetdiff(hes_aurum_primary2, hes_aurum_primary1, all = TRUE)
hes_aurum_primary = hes_aurum_primary1
nrow(hes_aurum_primary)

#comorb_aurum_hes[,database := "HES APC AURUM"] 
#comorb_aurum_hes[,codetype := "ICD"] 

setnames(hes_aurum_primary, old=c("ICD_PRIMARY", "ICDx"), new =c("diag_code", "icd_modifier")) # date of diagnosis = date of discharge 
nrow(hes_aurum_primary[is.na(discharged),]) # double check number of corrupt eventdates
hes_aurum_primary[, eventdate:= discharged ] # date of diagnosis = date of discharge 
hes_aurum_primary[is.na(eventdate), eventdate:= admidate ] # where discharge date is missing, use admission date
nrow(hes_aurum_primary[is.na(eventdate),]) # double check number of corrupt eventdates
hes_aurum_primary = hes_aurum_primary[!is.na(eventdate)] # exclude events where both admission and discharge dates are missing
nrow(hes_aurum_primary[is.na(eventdate)])

hes_aurum_primary = hes_aurum_primary[duplicated(hes_aurum_primary)==F,] # Exclude duplicates
format(nrow(hes_aurum_primary), big.mark="'") 
#comorb_aurum_hes[,diag_code_unique_ID := paste("I",diag_code, sep="_")] ## Define unique_ID
colnames(hes_aurum_primary)
hes_aurum_primary = hes_aurum_primary[,.(patid, eventdate, diag_code, admidate, discharged, spno, ICD_5, icd_modifier)]

# Merge with code dictionary
hes_aurum_primary2 = merge(hes_aurum_primary, icd_codes_table[disease_category == "cardiovascular",.(alt_icd_code, icd_code, unique_ID, description, disease, disease_category,disease_subcategory, disease_short, include)], by.x = "diag_code", by.y = "icd_code", all.x=T, allow.cartesian=TRUE)
levels(as.factor(fsetdiff(hes_aurum_primary[,.(diag_code)], hes_aurum_primary2[,.(diag_code)], all = TRUE)))
levels(as.factor(fsetdiff(hes_aurum_primary2[,.(diag_code)], hes_aurum_primary[,.(diag_code)], all = TRUE)))
#icd_codes_table[icd_code %in% c("I26.0", "I26.9", "I81", "I82.3")]
hes_aurum_primary = hes_aurum_primary2
setnames(hes_aurum_primary, old=c("alt_icd_code", "unique_ID"), new =c("alt_code", "diag_code_unique_ID")) # rename columns
colnames(hes_aurum_primary)

# Check that all codes have been matched with a disease
nrow(hes_aurum_primary[is.na(disease)])

# Define code type and database
hes_aurum_primary[,database := "HES AURUM PRIMARY"] 
hes_aurum_primary[,codetype := "icd"] 

colnames(hes_aurum_primary)
format(nrow(hes_aurum_primary), big.mark="'")  # number of records in table
format(length(unique(hes_aurum_primary$patid)), big.mark="'")  # number of unique patient IDs in table
```



### Combine all diagnoses in CPRD Aurum - Primary care, HES APC and HES OP
```{r combine_aurum, eval=TRUE, warning=FALSE, message=FALSE}
# Check
colnames(aurum_diagnoses_primary_care_dict)
class(aurum_diagnoses_primary_care_dict$patid)


# Rbind tables
include_cvd_death
if (include_cvd_death == "yes") {aurum_diagnoses_all = rbind( aurum_diagnoses_primary_care_dict, aurum_diagnoses_HES_APC_dict,aurum_diagnoses_HES_OP_dict, death_aurum_melted_cvd,  fill=TRUE ) }  else { aurum_diagnoses_all = rbind( aurum_diagnoses_primary_care_dict, aurum_diagnoses_HES_APC_dict, aurum_diagnoses_HES_OP_dict, fill=TRUE )} # if include_cvd_death == "yes" then include records of CVD death, else do not includes records of CVD death
#aurum_diagnoses_all = rbind(aurum_diagnoses_primary_care_dict, aurum_diagnoses_HES_APC_dict,aurum_diagnoses_HES_OP_dict, fill=TRUE )
nrow(aurum_diagnoses_all)

# Add pracid
aurum_diagnoses_all = merge(aurum_diagnoses_all, denominator_aurum[,.(patid, pracid)], by = "patid", all.x = TRUE)
summary(aurum_diagnoses_all$pracid)

# Check
# Check that for all diag_codes include == 1
levels(as.factor(aurum_diagnoses_all$include))

# Check eventdates and patid
nrow(aurum_diagnoses_all[is.na(eventdate),])
nrow(aurum_diagnoses_all[is.na(pracid),])
summary(aurum_diagnoses_all$pracid )

# define db_patid across gold and aurum
aurum_diagnoses_all[,db_patid := paste("A", patid, sep="")] 
```


## Save_1
```{r save_1, eval=TRUE, warning=FALSE, message=FALSE}
rm(list=setdiff(ls(), c("gold_diagnoses_all",  "aurum_diagnoses_all", "death_gold_melted_cvd", "death_aurum_melted_cvd", "hes_gold_primary", "hes_aurum_primary", "study_start_date", "study_end_date", "PATH", "PATH_D", "PATH_D2", "PATH_W", "PATH_W_orig", "PATH_L", "gold_codes_table", "aurum_codes_table", "icd_codes_table", "icd_9_10_codes_table", "opcs_codes_table", "all_codes_table", "denominator", "age_group_table", "compute_days", "age_group_fun", "icd_codes_to_be_double_checked_for_inclusion_gold", "diag_codes_table_inc_0", "gold_codes_to_be_double_checked_for_inclusion", "aurum_codes_to_be_double_checked_for_inclusion", "icd_codes_to_be_double_checked_for_inclusion_aurum", "include_cvd_death", "death_gold", "death_aurum")))
#load(file.path(PATH_W,"2022-02-18SelectedWorkspaceIdentification_1.RData"))
#save.image(file.path(PATH_W,"2022-02-18SelectedWorkspaceIdentification_1.RData"))
gc()
```


# Combine gold and aurum diagnoses
```{r, eval=TRUE, warning=FALSE, message=TRUE}
class(gold_diagnoses_all$patid)
class(aurum_diagnoses_all$patid)

gold_diagnoses_all[,patid:=as.integer64(patid)]
head(gold_diagnoses_all)
diagnoses_all = rbind(gold_diagnoses_all, aurum_diagnoses_all, fill=TRUE) # codes with include == 1 or 1-possible
nrow(diagnoses_all)
nrow(diagnoses_all[is.na(disease)])
head(diagnoses_all[is.na(disease)])
#levels(as.factor(diagnoses_all[is.na(disease)]$database))
nrow(diagnoses_all[is.na(pracid)])
nrow(diagnoses_all[is.na(eventdate)])

# DEATH All
death_gold[,db_patid := paste("G", patid, sep="")]
nrow(death_gold)
death_aurum[,db_patid := paste("A", patid, sep="")]
nrow(death_aurum)
death_all = rbind(death_gold, death_aurum, fill=TRUE)
nrow(death_all)

# DEATH (CVD only)
class(death_aurum_melted_cvd$patid)
class(death_gold_melted_cvd$patid)
death_gold_melted_cvd[,patid:=as.integer64(patid)]
death_gold_melted_cvd[,db_patid := paste("G", patid, sep="")]
death_aurum_melted_cvd[,db_patid := paste("A", patid, sep="")]
death_melted_cvd = rbind(death_gold_melted_cvd, death_aurum_melted_cvd, fill=TRUE)
nrow(death_melted_cvd)
nrow(death_melted_cvd[is.na(disease)])
nrow(death_melted_cvd[is.na(patid)])

# SENS (HES PRIMARY ONLY)
class(hes_gold_primary$patid)
class(hes_aurum_primary$patid)
hes_gold_primary[,patid:=as.integer64(patid)]
hes_gold_primary[,db_patid := paste("G", patid, sep="")]
hes_aurum_primary[,db_patid := paste("A", patid, sep="")]
hes_primary = rbind(hes_gold_primary, hes_aurum_primary, fill=TRUE)
nrow(hes_primary)
nrow(hes_primary[is.na(disease)])
nrow(hes_primary[is.na(patid)])
```


# Create code occurence tables - Any occurence
```{r code_occ, eval=TRUE, warning=FALSE, message=FALSE}
# Number of occurences by code with include == 1/1-possible (any occurence)
code_occ_1= diagnoses_all[,.N, by = c("diag_code_unique_ID","disease", "codetype" )]
code_occ_1[, "frequency":=paste(round(N/sum(N)*100,2),"%"), by=disease]
code_occ_1_dict = merge(code_occ_1[, .(disease, diag_code_unique_ID, N,frequency )], all_codes_table, by.x = c("disease", "diag_code_unique_ID"), by.y = c("disease", "unique_ID"), all = TRUE)
setcolorder(code_occ_1_dict, c("disease_category", "disease", "description", "N", "frequency","diag_code_unique_ID"))
code_occ_1_dict=code_occ_1_dict[order(disease, -N)]
nrow(code_occ_1_dict)
code_occ_1_dict[is.na(diag_code_unique_ID)]
levels(as.factor(diagnoses_all[is.na(diag_code_unique_ID)]$database))
diagnoses_all[is.na(diag_code_unique_ID)]

write.table(code_occ_1_dict, file.path(PATH,"/Phenotyping/Final_AD_INF_CVD/CVD_incidence/2023-03-11_code_occ_table_1_any_occurence.txt"), sep="\t", row.names=FALSE)

# Check number of unique patients identified by disease
#gold_diagnoses_all = gold_diagnoses_all[exclude == 0]
code_occ_table_by_disease =diagnoses_all[, .(count = uniqueN(patid)), by = c("disease", "disease_category")]
code_occ_table_by_disease=code_occ_table_by_disease[order(disease_category, disease)]
aggregate(code_occ_table_by_disease$count, by=list(Category=code_occ_table_by_disease$disease_category), FUN=sum)
code_occ_table_by_disease
write.table(code_occ_table_by_disease, file.path(PATH,"/Phenotyping/Final_AD_INF_CVD/CVD_incidence/2023-03-11_code_occ_table_by_disease_any_occurence.txt"), sep="\t")
```

# Save_2
```{r save_2, eval=TRUE, warning=FALSE, message=FALSE}
#save.image(file.path(PATH_W,"2022-02-18FullWorkspaceIdentification.RData"))
fwrite(diagnoses_all, file.path(PATH_W, "diagnoses_all.txt.gz"))
fwrite(death_melted_cvd, file.path(PATH_W, "death_melted_cvd.txt.gz"))
fwrite(hes_primary, file.path(PATH_W, "hes_primary.txt.gz"))
fwrite(death_all, file.path(PATH_W, "death_all.txt.gz"))


rm(list=setdiff(ls(), c("diagnoses_all", "death_melted_cvd", "hes_primary", "diagnoses_all_0", "study_start_date", "study_end_date", "PATH", "PATH_D", "PATH_D2", "PATH_W","PATH_W_orig", "PATH_L", "gold_codes_table", "aurum_codes_table", "icd_codes_table", "icd_9_10_codes_table", "opcs_codes_table", "all_codes_table", "denominator", "age_group_table", "compute_days", "age_group_fun", "icd_codes_to_be_double_checked_for_inclusion_gold", "diag_codes_table_inc_0", "gold_codes_to_be_double_checked_for_inclusion", "aurum_codes_to_be_double_checked_for_inclusion", "icd_codes_to_be_double_checked_for_inclusion_aurum", "include_cvd_death")))
#load(file.path(PATH_W,"2022-02-18SelectedWorkspaceIdentification_2.RData"))
save.image(file.path(PATH_W,"2023-03-03_SelectedWorkspaceIdentification_CVD_2.RData"))
#load(file.path(PATH_W,"2022-02-18SelectedWorkspaceIdentification_2.RData"))
gc()
```

# Select first date of diagnosis
## Select first date of diagnosis for each patient and EACH CONDITION
```{r first_date, eval=TRUE, warning=FALSE, message=FALSE}
start_time <- Sys.time()
diagnoses_all_first=diagnoses_all[,.SD[which.min(eventdate)], by = c("db_patid", "disease")]
end_time <- Sys.time()
end_time - start_time 

# Number of diagnoses
nrow(diagnoses_all_first)
class(diagnoses_all$eventdate)
```

# Save_3
```{r save_3, eval=TRUE, warning=FALSE, message=FALSE}
#save.image(file.path(PATH_W,"2022-02-18FullWorkspaceIdentification.RData"))
fwrite(diagnoses_all_first, file.path(PATH_W, "diagnoses_all_first.txt.gz"))
rm(list=setdiff(ls(), c("diagnoses_all_first", "diagnoses_all_AD_first", "diagnoses_all_CVD_first","study_start_date", "study_end_date", "PATH", "PATH_D", "PATH_D2", "PATH_W", "PATH_W_orig", "PATH_L", "gold_codes_table", "aurum_codes_table", "icd_codes_table", "icd_9_10_codes_table", "opcs_codes_table", "all_codes_table", "denominator", "cohort_1", "age_group_table", "compute_days", "age_group_fun", "icd_codes_to_be_double_checked_for_inclusion_gold", "diag_codes_table_inc_0", "gold_codes_to_be_double_checked_for_inclusion", "aurum_codes_to_be_double_checked_for_inclusion", "icd_codes_to_be_double_checked_for_inclusion_aurum", "include_cvd_death")))
#load(file.path(PATH_W,"2022-02-18SelectedWorkspaceIdentification_3.RData"))
#save.image(file.path(PATH_W,"2022-02-18SelectedWorkspaceIdentification_3.RData"))
gc()
```

# Create code occurence tables
```{r code_occ_1, eval=TRUE, warning=FALSE, message=FALSE}
# Number of occurences by code with include == 1 / 1-possible (any first occurence)
code_occ_1= diagnoses_all_first[,.N, by = c("diag_code_unique_ID","disease", "codetype" )]
code_occ_1[, "frequency":=paste(round(N/sum(N)*100,2),"%"), by=disease]
code_occ_1_dict = merge(code_occ_1[, .(disease, diag_code_unique_ID, N,frequency )], all_codes_table, by.x = c("disease", "diag_code_unique_ID"), by.y = c("disease", "unique_ID"), all = TRUE)
setcolorder(code_occ_1_dict, c("disease_category", "disease", "description", "N", "frequency","diag_code_unique_ID"))
code_occ_1_dict=code_occ_1_dict[order(disease, -N)]
nrow(code_occ_1_dict)
write.table(code_occ_1_dict, file.path(PATH,"/Phenotyping/Final_AD_INF_CVD/CVD_incidence/Code_occurence_tables_any_time/2023-03-11_code_occ_table_1_first.txt"), sep="\t", row.names=FALSE)

# Number of occurences all codes including "0 - possible" codes
code_occ_0_dict = merge(code_occ_1[, .(disease, diag_code_unique_ID, N,frequency )], diag_codes_table_inc_0, by.x = c("disease", "diag_code_unique_ID"), by.y = c("disease", "unique_ID"), all = TRUE)
setcolorder(code_occ_0_dict, c("disease_category", "disease", "description", "N", "frequency","diag_code_unique_ID"))
code_occ_0_dict=code_occ_0_dict[order(disease, -N)]
nrow(code_occ_0_dict)
code_occ_0_dict

# Codes to be double-checked
icd_codes_to_be_double_checked_for_inclusion = rbind(icd_codes_to_be_double_checked_for_inclusion_gold, icd_codes_to_be_double_checked_for_inclusion_aurum)
icd_codes_to_be_double_checked_for_inclusion = icd_codes_to_be_double_checked_for_inclusion[,.(count_na=sum(count_na)),.(disease_category, disease, diag_code, description, unique_ID)]
icd_codes_to_be_double_checked_for_inclusion[,codetype := "icd"]
gold_codes_to_be_double_checked_for_inclusion[,codetype := "gold"]
aurum_codes_to_be_double_checked_for_inclusion[,codetype := "aurum"]
aurum_codes_to_be_double_checked_for_inclusion[, diag_code_aurum:= as.character(diag_code_aurum)]
codes_to_be_double_checked_for_inclusion = rbind(icd_codes_to_be_double_checked_for_inclusion, gold_codes_to_be_double_checked_for_inclusion, aurum_codes_to_be_double_checked_for_inclusion, fill=TRUE)
head(codes_to_be_double_checked_for_inclusion)
setcolorder(codes_to_be_double_checked_for_inclusion, c("disease_category", "disease", "description", "count_na","unique_ID", "diag_code", "diag_code_aurum"))
codes_to_be_double_checked_for_inclusion=codes_to_be_double_checked_for_inclusion[order(disease_category, disease, -count_na)]
codes_to_be_double_checked_for_inclusion
write.table(codes_to_be_double_checked_for_inclusion, file.path(PATH,"/Phenotyping/Final_AD_INF_CVD/CVD_incidence/Code_occurence_tables_any_time/2023-03-11_codes_to_be_double_checked_for_inclusion.txt"), sep="\t") # these are "0-possible" codes (ie. codes excluded from analysis but originally one of the reviewer sugggested to include that code) alongside number of occurences

# Check number of unique patients identified by disease
#gold_diagnoses_all = gold_diagnoses_all[exclude == 0]
code_occ_table_by_disease =diagnoses_all_first[, .(count = uniqueN(patid)), by = c("disease", "disease_category")]
code_occ_table_by_disease=code_occ_table_by_disease[order(disease_category, disease)]
aggregate(code_occ_table_by_disease$count, by=list(Category=code_occ_table_by_disease$disease_category), FUN=sum)
code_occ_table_by_disease
write.table(code_occ_table_by_disease, file.path(PATH,"/Phenotyping/Final_AD_INF_CVD/CVD_incidence/Code_occurence_tables_any_time/2023-03-11_code_occ_table_by_disease.txt"), sep="\t")
```

# Save_3b
```{r save_3b, eval=TRUE, warning=FALSE, message=FALSE}
#save.image(file.path(PATH_W,"2022-02-18FullWorkspaceIdentification.RData"))
#fwrite(diagnoses_all_first, file.path(PATH_W, "diagnoses_all_first.txt.gz"))
rm(list=setdiff(ls(), c("diagnoses_all_first", "diagnoses_all_AD_first", "diagnoses_all_CVD_first","study_start_date", "study_end_date", "PATH", "PATH_D", "PATH_D2", "PATH_W", "PATH_W_orig","PATH_L", "gold_codes_table", "aurum_codes_table", "icd_codes_table", "icd_9_10_codes_table", "opcs_codes_table", "all_codes_table", "denominator", "cohort_1", "age_group_table", "compute_days", "age_group_fun", "icd_codes_to_be_double_checked_for_inclusion_gold", "diag_codes_table_inc_0", "gold_codes_to_be_double_checked_for_inclusion", "aurum_codes_to_be_double_checked_for_inclusion", "icd_codes_to_be_double_checked_for_inclusion_aurum", "include_cvd_death")))
#load(file.path(PATH_W,"2022-02-18SelectedWorkspaceIdentification_3b.RData"))
save.image(file.path(PATH_W,"2023-03-03_SelectedWorkspaceIdentification_CVD_3b.RData"))
#load(file.path(PATH_W,"2022-02-18SelectedWorkspaceIdentification_3b.RData"))
gc()
```




# Save_5
```{r save_M, eval=TRUE, warning=FALSE, message=FALSE}
#save.image(file.path(PATH_W,"2022-02-18FullWorkspaceIdentification.RData"))
rm(list=setdiff(ls(), c("diagnoses_all_first_denom", "matching_cohort_no_CVD", "study_start_date", "study_end_date", "PATH", "PATH_D", "PATH_D2", "PATH_W","PATH_W_orig", "PATH_L", "gold_codes_table", "aurum_codes_table", "icd_codes_table", "icd_9_10_codes_table", "opcs_codes_table", "all_codes_table", "denominator", "cohort_1", "age_group_table", "compute_days", "age_group_fun", "icd_codes_to_be_double_checked_for_inclusion_gold", "diag_codes_table_inc_0", "gold_codes_to_be_double_checked_for_inclusion", "aurum_codes_to_be_double_checked_for_inclusion", "icd_codes_to_be_double_checked_for_inclusion_aurum", "include_cvd_death")))

save.image(file.path(PATH_W,"2023-03-03_SelectedWorkspaceIdentification_CVD_5.RData"))
```
```